
10-100 kb windows in 500kb invs is 50 windows per inv
Invs larger than 1.8kb have >5% of the inversion covered by exclusive windows
78 inversions are at least 10kb

I want to 
1. Make sliding windows for repeat regions,  Take only those that overlap with the workspace (with bedtools makewindows and intersect)


```{bash}
TODAY="defTest"
DIR="analysis/${TODAY}_LocationPatterns"
CONDITIONS="avgBherer CEUSpence"
OUTDIR="${DIR}/slWins/"
HEADER="Chromosome\tStart\tEnd\tChrRegion"
mkdir -p $OUTDIR

for CONDITION in $CONDITIONS; do

  ISOCHORES="${DIR}/GATfiles/Isochores/isochores_4breaks_${CONDITION}.bed"
  WORKSPACE="${DIR}/GATfiles/Workspace/workspace_${CONDITION}.bed"
  WINDOW="$OUTDIR/slWins_${CONDITION}.bed"
  FILWINDOW="$OUTDIR/slWins_${CONDITION}_filtered.bed"
  
  bedtools makewindows -b $ISOCHORES -w 100000 -s 10000 -i src > $WINDOW
  
  echo -e $HEADER > $FILWINDOW
  bedtools intersect -a $WINDOW -b $WORKSPACE -u >> $FILWINDOW
  
done
```

1.1 For each window make max, min, weighted RecRate (with fillWindowMeasurements)

```{bash}
TODAY="defTest"
DIR="analysis/${TODAY}_LocationPatterns"
#CONDITIONS="avgBherer CEUSpence"

ENVPY="/home/rgomez/anaconda3/bin/python"

#Bherer
CONDITION="avgBherer"
MAP="data/Bherer_Refined_genetic_map_b37_processed/sexavg_allChr.bed"
FILWINDOW="$DIR/slWins/slWins_${CONDITION}_filtered.bed"

$ENVPY code/python/fillWindowMeasurements.py -m $MAP -w $FILWINDOW -o $DIR/slWins -d $CONDITION


#Bherer
CONDITION="CEUSpence"
MAP="data/SpenceSong_hg19_recMaps_processed/CEU_recombination_map_hg19_allChr.bed"
FILWINDOW="$DIR/slWins/slWins_${CONDITION}_filtered.bed"

$ENVPY code/python/fillWindowMeasurements.py -m $MAP -w $FILWINDOW -o $DIR/slWins -d $CONDITION


```


2. Take windows overlapping inversions (segments) as observed


```{bash}
TODAY="defTest"
DIR="analysis/${TODAY}_LocationPatterns/"
CONDITIONS="avgBherer CEUSpence"
OUTDIR="${DIR}/slWins/"
#HEADER="Chromosome\tStart\tEnd\tChrRegion"
mkdir -p $OUTDIR

for CONDITION in $CONDITIONS; do

   WINDOW="$OUTDIR/slWins_${CONDITION}.bed"
   INVCOORD="${DIR}/GATfiles/Segments_byInv/*"
  INVWINDOW="$OUTDIR/slWins_${CONDITION}_invs.bed"
  
  bedtools intersect -wa -wb -a $WINDOW -b $INVCOORD   > $INVWINDOW
  
done
```


3. Compare - THIS IS A FUNCTION
I can use it later

```{r}
library(ggplot2)
condition<-"avgBherer"

# Upload genome measurements for a given condition
genome<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/fillWindowMeasurements/",condition,"/windowData.txt"), header = T, stringsAsFactors = F)

# Upload inversion windows
invs<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/slWins_",condition,"_invs.bed"), stringsAsFactors = F)
colnames(invs)<-c("Chromosome", "Start", "End", "ChrRegion", "order", "invChromosome", "invStart", "invEnd", "inv")
```


```{r function_sampleGenome}
sampleGenome<-function(invs, genome, repDens, nSamples){
    
    # For a repeat density, select corresponding inversions
    whichRep<-data.frame(table( invs$inv, invs$ChrRegion))
    colnames(whichRep)<-c("inv", "ChrRegion", "Freq")
    whichRep<-whichRep[whichRep$Freq >0,]
    
    heteroRep<-whichRep[duplicated(whichRep$inv), "inv"]
    
    subRep<-whichRep[!(whichRep$inv %in% heteroRep) & whichRep$ChrRegion %in% repDens,"inv" ]
    
    # Take info for selected rep density
    genoInfo<-genome[genome$ChrRegion %in%  repDens,]
    rownames(genoInfo)<-paste(genoInfo$Chromosome, genoInfo$Start, genoInfo$End, sep = "_")
    
    # Replace info for inv windows
    invInfo<-invs[invs$inv %in% as.character(subRep),]
    invInfo$tags<-paste(invInfo$Chromosome, invInfo$Start, invInfo$End, sep = "_")
    
    genoInfo[invInfo$tags, "ChrRegion" ]<-invInfo$inv
    
    ###MAK ELOOP
    allinvsamples<-data.frame()
    for (inversion in subRep){
      
      sampleSize<-nrow(invInfo[invInfo$inv == inversion,])
      
      # Make subset  
      toSam<-genoInfo[genoInfo$ChrRegion %in%  repDens,]
      toSam$previous<-toSam$Start - c(0, toSam$Start[1:length(toSam$Start)-1])
      
      # Make sure it is sorted
      toSam<-toSam[order(toSam$Chromosome, toSam$Start),]
      
      # Give it rownames
      rownames(toSam)<-c(1:nrow(toSam))
      
      # Select sampleable rownames
      nonums<-c()
      nostarters<-as.numeric(rownames(toSam[toSam$previous != 10000,]))-sampleSize+1
      for(n in nostarters){
        nonums<-c(nonums, c(n:(n+sampleSize-2)))
      }
      
      options<-as.numeric(rownames(toSam)[!(as.numeric(rownames(toSam)) %in% nonums)])
  
      # Make random samples
      samples<-data.frame()
      
      starters<-sample(options, nSamples, replace = F)
      samples<-data.frame()
        for(i in starters){
          sample<-toSam[seq(i,i+sampleSize-1),]
          sample$name<-i
          samples<-rbind(samples,sample)
        }
      samples$inv<-inversion
      allinvsamples<-rbind(allinvsamples, samples)
    }
    
return(list(allinvsamples,genoInfo[!(genoInfo$ChrRegion %in%  repDens),] ))
}
```


```{r}
# Test function
repDens<-"High005"
nSamples<-1000

samples<-sampleGenome(invs=invs, genome=genome, repDens=repDens, nSamples=nSamples)


```



I can also make a meta-analysis of:
min and max weighted recRate for inv vs the same for an equivalent number of windows randomly selected

OK BUT
what they comapre is the ratio between population and pedigree, so I need the same windows for each...
1.1 For each window make max, min, weighted RecRate (with fillWindowMeasurements), FOR THE OTHER WINDOW SET (I make both just in case one of them doesn't work)

```{bash}
TODAY="defTest"
DIR="analysis/${TODAY}_LocationPatterns/"

#CONDITIONS="avgBherer CEUSpence"

ENVPY="/home/rgomez/anaconda3/bin/python"

#Bherer
CONDITION="CEUSpence"
MAP="data/Bherer_Refined_genetic_map_b37_processed/sexavg_allChr.bed"
FILWINDOW="$DIR/slWins/slWins_${CONDITION}_filtered.bed"

$ENVPY code/python/fillWindowMeasurements.py -m $MAP -w $FILWINDOW -o $DIR/slWins/flipped -d $CONDITION


#Bherer
CONDITION="avgBherer"
MAP="data/SpenceSong_hg19_recMaps_processed/CEU_recombination_map_hg19_allChr.bed"
FILWINDOW="$DIR/slWins/slWins_${CONDITION}_filtered.bed"

$ENVPY code/python/fillWindowMeasurements.py -m $MAP -w $FILWINDOW -o $DIR/slWins/flipped -d $CONDITION


```

Now I'll use the avgBherer with Spence map which is the one that worked best:

```{r}
library(ggplot2)
condition<-"avgBherer"
repDens<-"High005"

# Upload genome measurements for a given condition
genome_ped<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/fillWindowMeasurements/",condition,"/windowData.txt"), header = T, stringsAsFactors = F)
genome_pop<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/flipped/fillWindowMeasurements/",condition,"/windowData.txt"), header = T, stringsAsFactors = F)

# Ratio of the two RATES ped/pop
genome<-merge(genome_ped, genome_pop, by = c("Chromosome", "Start", "End", "ChrRegion"), suffixes = c("_ped", "_pop") )
genome$WAvgRate_ped_pop<-genome$WAvgRate_ped/genome$WAvgRate_pop

# Upload inversion info
invs<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/slWins_",condition,"_invs.bed"), stringsAsFactors = F)
colnames(invs)<-c("Chromosome", "Start", "End", "ChrRegion", "order", "invChromosome", "invStart", "invEnd", "inv")

#Make samples
samples<-sampleGenome(invs=invs, genome=genome, repDens =repDens, nSamples = 1000)
sampleMeasurements<-samples[[1]]
invMeasurements<-samples[[2]]

mins_dist<-aggregate( WAvgRate_ped_pop ~ name+inv, sampleMeasurements, min)
maxs_dist<-aggregate( WAvgRate_ped_pop ~ name+inv, sampleMeasurements, max)
randist<-merge(mins_dist, maxs_dist, by = c("name", "inv"), suffixes = c(".min", ".max"))
randist<-melt(randist, id.vars = c("inv", "name"))

mins_ori<-aggregate(WAvgRate_ped_pop ~ ChrRegion, invMeasurements, min )
maxs_ori<-aggregate(WAvgRate_ped_pop ~ ChrRegion, invMeasurements, max )
oridist<-merge(mins_ori, maxs_ori, by = c( "ChrRegion"), suffixes = c(".min", ".max"))
oridist<-melt(oridist, id.vars = c("ChrRegion"))
colnames(oridist)<-c("inv", "variable", "value")

ggplot()+geom_boxplot(data=randist, aes(x = inv, y = value))+
  geom_point(data=oridist, aes(x = inv, y = value), color = "red")+
  scale_y_log10()+
  facet_grid(variable~.)

```


SHOULD REPEAT THOSE ANALYSES THAT DEPEND ON FILLWINDOWMEASUREMENTS!

# Invented method

## Step 1
I will need: sliding windows with min, max, weighted recRate (done)

I want to 
1. Make sliding windows for the genome,  Take only those that overlap with the workspace (with bedtools makewindows and intersect)
I use Bherer to make the windows, because we wnat shared windows between both maps

```{bash}
TODAY="defTest"
DIR="analysis/${TODAY}_LocationPatterns"
CONDITION="avgBherer"
OUTDIR="${DIR}/slWins/"
HEADER="Chromosome\tStart\tEnd"
mkdir -p $OUTDIR
WINNAME="100k_10k"

#MAke windows
  ISOCHORES="${DIR}/GATfiles/Workspace/workspace_${CONDITION}.bed"
  WORKSPACE="${DIR}/GATfiles/Workspace/workspace_${CONDITION}.bed"
  WINDOW="$OUTDIR/slWins_${WINNAME}_${CONDITION}.bed"
  FILWINDOW="$OUTDIR/slWins_${WINNAME}_${CONDITION}_filtered.bed"


  echo -e $HEADER > $WINDOW
  bedtools makewindows -b $ISOCHORES -w 100000 -s 10000  > $WINDOW
  
  #Take only those that overlap with workspace
 # echo -e $HEADER > $FILWINDOW
#  bedtools intersect -a $WINDOW -b $WORKSPACE -u >> $FILWINDOW
```


```{bash}
# 1.1 For each window make  fillWindowMeasurements

ENVPY="/home/rgomez/anaconda3/bin/python"

#Bherer

MAP="data/Bherer_Refined_genetic_map_b37_processed/sexavg_allChr.bed"
$ENVPY code/python/fillWindowMeasurements_avgonly.py -m $MAP -w $FILWINDOW -o $DIR/slWins/flipped -d "${WINNAME}_${CONDITION}wins_inBherermap"

#Spence

MAP="data/SpenceSong_hg19_recMaps_processed/CEU_recombination_map_hg19_allChr.bed"
$ENVPY code/python/fillWindowMeasurements_avgonly.py -m $MAP -w $FILWINDOW -o $DIR/slWins/flipped  -d "${WINNAME}_${CONDITION}wins_inSpencemap"


```




```{r}
library(ggplot2)
condition<-"avgBherer"

# Upload genome measurements for a given condition
genome<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/fillWindowMeasurements/",condition,"/windowData.txt"), header = T, stringsAsFactors = F)

```

These sliding windows will be used to compare rec measurements inside the actual and simulated inversions.
## Step 2 Generate inv coordinate tables (actual and simulated)
OJO! 

Why do I sample the genome and not all the possible numbers? The only reason is because a vector with all the positions in the genome would be very large, while sampling windows is a good approximation. I exclude those windows overlapping with inversions and I leave space at the start and the end of the chromosome to not cut the inversion, but I do not leave space for flanking regions because it could happen in real life that I do not have enough space for flanking regions of a determinate size. 

> In addition, I want to make sure that permutations do not overlap other invesions, even if I'm looking only at big invs. This is somethong that Basatena does to avoid biases. 

```{r}
condition<-"avgBherer"
# These are INCLUSIVE made with bedtools
invs<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/slWins_",condition,"_invs.bed"), stringsAsFactors = F)
colnames(invs)<-c("Chromosome", "Start", "End", "ChrRegion", "order", "invChromosome", "invStart", "invEnd", "inv")
invs$invSize<-invs$invEnd - invs$invStart
```

```{r}
# OJO! measurements overlap a lot for invs < 100 kb (sliding window size). They use originally this size for invs > 500kb
# I will make a """general""" dataset (> 100kb) and a "paper equivalent" dataset (> 500kb)
#This will make simulations go faster
sort(unique(invs[invs$inv %in% biginvs, "invSize"]))
biginvs<-unique(invs[invs$invSize >= 50000, "inv"])
verybiginvs<-unique(invs[invs$invSize >= 500000, "inv"])

tail(unique(invs[order(invs$invSize),c("inv", "invSize")]), 20)

```

The max samples we can make with "same" are 3992 (total available positions), and are very fragmented, so big inversions do not fit. Big inversions need to be simulated in all the genome.

```{r function_sampleGenome}
dir.create( "analysis/defTest_LocationPatterns/slWins/adjacentCoords/")
sampleGenome<-function(toSample, invInfo, genoInfo, repDens, nSamples, store){
  slidingSize<-10000
  ####################################################
    # toSample is the list of invs we want to simulate
    # invInfo is the list of all invs (we want to avoid overlaps with these)
    # genome is the dataset to sample from
    # repDens can be "same" to resample in the main repDens profile of the real inv or "all" to resample in all the genome
    # nSamples is sample size
  # invInfo <- invs
  # genoInfo<-genome
  # repDens<-"same"
  # nSamples<-5000
  ####################################################
  
    # TAKE INVERSION SAMPLE TEMPLATES
      toSampleInfo<-invInfo[invInfo$inv %in% toSample,]
    
    # DETERMINE REPEAT STATUS for sample templates
      repeatsTable<-data.frame(table(toSampleInfo$inv, toSampleInfo$ChrRegion))
      repeatsTable<-merge(aggregate(Freq ~ Var1, data = repeatsTable, FUN = "max"), repeatsTable, all.x = T)
      
    # GENOME FILTERIGNG
      # Tag genome
      rownames(genoInfo)<-paste(genoInfo$Chromosome, genoInfo$Start, genoInfo$End, sep = "_")
     
      # Tag inv list
      invInfo$tags<-paste(invInfo$Chromosome, invInfo$Start, invInfo$End, sep = "_")
      
      # Replace Repeat info to avoid ALL inv regions
      genoInfo[invInfo$tags, "ChrRegion" ]<-invInfo$inv
      
      # Now, for each used repeat status OR for all the genome, we want to make a list of all the sampleable sites
      sampleables<-list()
      if(repDens == "same"){
        for (repD in unique(repeatsTable$Var2)){
         sampleables[[repD]]<-genoInfo[genoInfo$ChrRegion == repD,] 
        }
      }else{
        sampleables[["all"]]<-genoInfo[grep("HsInv" , genoInfo$ChrRegion, invert = T) ,]
      }
    
    # MAKE LOOP
    allinvsamples<-data.frame()

    for (inversion in toSample){
      
      # Calculate number of windows
      sampleSize<-nrow(invInfo[invInfo$inv == inversion,])
      
      # Select subset and prepare to calculate excluded coords  
      whereToSample<-ifelse(repDens == "all", "all", as.character(repeatsTable[repeatsTable$Var1 == inversion,"Var2"]))
      toSam<-sampleables[[whereToSample]]    
      toSam<-toSam[order(toSam$Chromosome, toSam$Start),]# sort
      toSam$previous<-toSam$Start - c(0, toSam$Start[1:length(toSam$Start)-1]) # mark previous coord
      rownames(toSam)<-c(1:nrow(toSam)) # give rownames
      
      # Select sampleable rownames
      nonums<-c()
      nostarters<-c(as.numeric(rownames(toSam[toSam$previous != slidingSize,])), nrow(toSam)+1)-sampleSize+1
      for(n in nostarters){
        nonums<-c(nonums, c(n:(n+sampleSize-2)))
      }
      
      options<-as.numeric(rownames(toSam)[!(as.numeric(rownames(toSam)) %in% nonums)])
  
      # Make random samples
      starters<-sample(options, nSamples, replace = F)
      
      # Generate a table like the invInfo table
      samples<-toSam[starters, c("Chromosome", "Start", "End", "ChrRegion")]
      samples$End<-samples$Start + unique(invInfo[invInfo$inv == inversion,"invSize"])
      samples$inv<-paste(inversion , c(1:nrow(samples)), sep = "_")
      
      
      # Add to table
      allinvsamples<-rbind(allinvsamples,samples)
    }
    colnames(allinvsamples)<-c("Chromosome", "invStart", "invEnd", "ChrRegion", "inv")
    write.table(allinvsamples, store, quote = F, sep = "\t", row.names = F, col.names = F)

return(allinvsamples)
}
```

```{r}
#Bigincs
biginvs_permutations_all<-sampleGenome(biginvs, invs, genome, "all", 10000, "analysis/defTest_LocationPatterns/slWins/adjacentCoords/in_biginvs_permutations_all.bed" )
# biginvs_permutations_same<-sampleGenome(biginvs, invs, genome, "same", 10000)

# Verybiginvs
verybiginvs_permutations_all<-sampleGenome(verybiginvs, invs, genome, "all", 10000,"analysis/defTest_LocationPatterns/slWins/adjacentCoords/in_verybiginvs_permutations_all.bed" )
# verybiginvs_permutations_same<-sampleGenome(verybiginvs, invs, genome, "same", 10000)


```


## Step 3 Core method (take measurements of avg, min, max fold change)

Since I will need permutations to compare to, all the following method should be grouped in easy-use functions that take as input inversions coordinates (actual or simulated). 
 I will also need the measurements for the windows outside, in fixed and proportional modes. In Seich Basatena paper they use fixed 0.75Mb = 750kb for a 0.75Mb simulated inv. I could use something near the median inv length, maybe (not the average bc there are very big invs). 

```{bash}
# args[1]<-"analysis/defTest_LocationPatterns/GATfiles/Segments_byInv/HsInv0501.bed" #Bedfile with invs to analyze
# args[2]<- 200000 #Confidence interval
# args[3]<-1000 # Number of samples
# args[4]<-"analysis/defTest_LocationPatterns/slWins/fillWindowMeasurements/avgBherer/windowData.txt" # POP map in sliding windows made without repeat regions
# args[5]<-"analysis/defTest_LocationPatterns/slWins/flipped/fillWindowMeasurements/avgBherer/windowData.txt" # PED map in sliding windows made without repeat regions

pwd
#"code/rscript/checkMapValidity.R"
```

```{bash}

INVLIST="HsInv1117 HsInv1065 HsInv0340 HsInv0370 HsInv0786 HsInv0553 HsInv1051 HsInv0573 HsInv0371 HsInv0379 HsInv0228 HsInv0382 HsInv1110 HsInv1057 HsInv0290 HsInv0486 HsInv0286 HsInv1111 HsInv0501"

for i in $INVLIST; do 
echo $i
Rscript code/rscript/checkMapValidity.R "analysis/defTest_LocationPatterns/GATfiles/Segments_byInv/${i}.bed" 200000 1000 "analysis/defTest_LocationPatterns/slWins/flipped/fillWindowMeasurements/100k_10k_avgBhererwins_inBherermap/windowData.txt" "analysis/defTest_LocationPatterns/slWins/flipped/fillWindowMeasurements/100k_10k_avgBhererwins_inSpencemap/windowData.txt" "tmp/${i}.txt"

done

```




```{r}

files<-grep("HsInv",list.files("tmp/"), value = T)

table<-data.frame()
for (i in files){
   table<-rbind(table, read.table(paste0("tmp/", i), header = T))
}

minmax<-unique(table[,c("inv", "minFC_pval", "maxFC_pval")])
mypval<-function( threshold, type){
  samples<-nrow(minmax)
  
  # If a max has a p-value >= 0.95, it has at must 5% of the distribution > observed
  #If a min has a p-value <= 0.05, it has at must 5% of the distribution < observed
  success<-ifelse(type =="max", sum(minmax$maxFC_pval >= 1-threshold), sum(minmax$minFC_pval <= threshold)) 
  
  pbinom(success-1, size = samples, prob = threshold, lower.tail = F) #Probabilidad de n o mas successes (= probabilidad de mas de n-1)

  }
    
  minmax
threshold=0.05
# The max
mypval(0.05, "max")
mypval(0.01, "max")
#The min
mypval(0.05, "min")
mypval(0.01, "min")

##Also, with my method regarding the surroundings
# sides<-table[table$variable == "WAvgRate_max",c("inv","Fold_change_pval_ped", "Fold_change_pval_pop")]
# sides<-reshape2::melt(sides)
# ggplot(sides)+geom_boxplot(aes(x = variable, y = value))+geom_jitter(aes(x = variable, y = value), width = 0.2)

## mockup 
minmax<-merge(minmax, unique(invs[,c("inv", "invSize")]))
minmax["Inversion length (Mb)"]<-round(minmax$invSize/1000000, 2)
minmax["% of empirical dist. < min ped/pop ratio"]<-minmax$minFC_pval*100
minmax["% of empirical dist. > max ped/pop ratio"]<-minmax$maxFC_pval<-(1-minmax$maxFC_pval)*100

minmax
print(minmax[order(minmax$`Inversion length (Mb)`, decreasing = T),c("inv", "Inversion length (Mb)", "% of empirical dist. > max ped/pop ratio", "% of empirical dist. < min ped/pop ratio")])

invs$worked<-invs$inv %in% minmax$inv
invlist<-unique(invs[, c("inv", "invSize", "worked")])
```

```{r}
library(ggplot2)
ggplot(table)+geom_boxplot(aes(x = variable, y = Fold_change_pval_ped))

library(vcfR)
vcf<-read.vcfR("data/InversionGenotypes_processed/allgenotypes.vcf")
vcfInfo<-vcfR2tidy(vcf, info_only = TRUE)$fix

freqs<-vcfInfo[,c("ID", "AF_EUR", "AF")]
freqs<-merge(freqs, table, by.x = "ID", by.y = "inv")

library(ggpubr)
a<-ggplot(freqs[freqs$variable == "WAvgRate_min",], aes(x = as.numeric(AF_EUR), y = Fold_change_pval_ped))+geom_point()+geom_smooth(method = "lm")+stat_cor(method = "pearson", label.x = 0.4)+ggtitle("Pedigree-based map")

b<-ggplot(freqs[freqs$variable == "WAvgRate_min",], aes(x = as.numeric(AF_EUR), y = Fold_change_pval_pop))+geom_point()+geom_smooth(method = "lm")+stat_cor(method = "pearson", label.x = 0.4)+ggtitle("Population-based map")

library(gridExtra)

grid.arrange(a,b)

toplot<-freqs[freqs$variable == "WAvgRate_min",c("AF_EUR", "Fold_change_pval_ped", "Fold_change_pval_pop")]
library(reshape2)
toplot<-melt(toplot)
# levels(toplot$variable)<-c("Pedigree-based map", "Population-based map")

ggplot(toplot, aes(x = as.numeric(AF_EUR), y = value, color =variable , fill = variable))+geom_point()+geom_smooth(method = "lm", alpha = 0.1 )+stat_cor(method = "pearson", label.x = 0.4)+ggtitle("Influence of frequency in CO rate reduction")+labs(x = "Frequency in europeans", y = "p-value of min rec.rate inside/outside ratio")


toplot<-freqs[freqs$variable == "WAvgRate",c("AF_EUR", "Fold_change_pval_ped", "Fold_change_pval_pop")]
toplot<-melt(toplot)

ggplot(toplot, aes(x = as.numeric(AF_EUR), y = value, color =variable , fill = variable))+geom_point()+geom_smooth(method = "lm", alpha = 0.1 )+stat_cor(method = "pearson", label.x = 0.4)+ggtitle("Influence of frequency in CO rate reduction")+labs(x = "Frequency in europeans", y = "p-value of average rec.rate inside/outside ratio")


toplot<-freqs[freqs$variable == "WAvgRate_max",c("AF_EUR", "Fold_change_pval_ped", "Fold_change_pval_pop")]
toplot<-melt(toplot)

ggplot(toplot, aes(x = as.numeric(AF_EUR), y = value, color =variable , fill = variable))+geom_point()+geom_smooth(method = "lm", alpha = 0.1 )+stat_cor(method = "pearson", label.x = 0.4)+ggtitle("Influence of frequency in CO rate reduction")+labs(x = "Frequency in europeans", y = "p-value of max rec.rate inside/outside ratio")


```

