
10-100 kb windows in 500kb invs is 50 windows per inv
Invs larger than 1.8kb have >5% of the inversion covered by exclusive windows
78 inversions are at least 10kb

I want to 
1. Make sliding windows for repeat regions,  Take only those that overlap with the workspace (with bedtools makewindows and intersect)


```{bash}
TODAY="defTest"
DIR="analysis/${TODAY}_LocationPatterns"
CONDITIONS="avgBherer CEUSpence"
OUTDIR="${DIR}/slWins/"
HEADER="Chromosome\tStart\tEnd\tChrRegion"
mkdir -p $OUTDIR

for CONDITION in $CONDITIONS; do

  ISOCHORES="${DIR}/GATfiles/Isochores/isochores_4breaks_${CONDITION}.bed"
  WORKSPACE="${DIR}/GATfiles/Workspace/workspace_${CONDITION}.bed"
  WINDOW="$OUTDIR/slWins_${CONDITION}.bed"
  FILWINDOW="$OUTDIR/slWins_${CONDITION}_filtered.bed"
  
  bedtools makewindows -b $ISOCHORES -w 100000 -s 10000 -i src > $WINDOW
  
  echo -e $HEADER > $FILWINDOW
  bedtools intersect -a $WINDOW -b $WORKSPACE -u >> $FILWINDOW
  
done
```

1.1 For each window make max, min, weighted RecRate (with fillWindowMeasurements)

```{bash}
TODAY="defTest"
DIR="analysis/${TODAY}_LocationPatterns"
#CONDITIONS="avgBherer CEUSpence"

ENVPY="/home/rgomez/anaconda3/bin/python"

#Bherer
CONDITION="avgBherer"
MAP="data/Bherer_Refined_genetic_map_b37_processed/sexavg_allChr.bed"
FILWINDOW="$DIR/slWins/slWins_${CONDITION}_filtered.bed"

$ENVPY code/python/fillWindowMeasurements.py -m $MAP -w $FILWINDOW -o $DIR/slWins -d $CONDITION


#Bherer
CONDITION="CEUSpence"
MAP="data/SpenceSong_hg19_recMaps_processed/CEU_recombination_map_hg19_allChr.bed"
FILWINDOW="$DIR/slWins/slWins_${CONDITION}_filtered.bed"

$ENVPY code/python/fillWindowMeasurements.py -m $MAP -w $FILWINDOW -o $DIR/slWins -d $CONDITION


```


2. Take windows overlapping inversions (segments) as observed


```{bash}
TODAY="defTest"
DIR="analysis/${TODAY}_LocationPatterns/"
CONDITIONS="avgBherer CEUSpence"
OUTDIR="${DIR}/slWins/"
#HEADER="Chromosome\tStart\tEnd\tChrRegion"
mkdir -p $OUTDIR

for CONDITION in $CONDITIONS; do

   WINDOW="$OUTDIR/slWins_${CONDITION}.bed"
   INVCOORD="${DIR}/GATfiles/Segments_byInv/*"
  INVWINDOW="$OUTDIR/slWins_${CONDITION}_invs.bed"
  
  bedtools intersect -wa -wb -a $WINDOW -b $INVCOORD   > $INVWINDOW
  
done
```


3. Compare - THIS IS A FUNCTION
I can use it later

```{r}
library(ggplot2)
condition<-"avgBherer"

# Upload genome measurements for a given condition
genome<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/fillWindowMeasurements/",condition,"/windowData.txt"), header = T, stringsAsFactors = F)

# Upload inversion windows
invs<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/slWins_",condition,"_invs.bed"), stringsAsFactors = F)
colnames(invs)<-c("Chromosome", "Start", "End", "ChrRegion", "order", "invChromosome", "invStart", "invEnd", "inv")
```


```{r function_sampleGenome}
sampleGenome<-function(invs, genome, repDens, nSamples){
    
    # For a repeat density, select corresponding inversions
    whichRep<-data.frame(table( invs$inv, invs$ChrRegion))
    colnames(whichRep)<-c("inv", "ChrRegion", "Freq")
    whichRep<-whichRep[whichRep$Freq >0,]
    
    heteroRep<-whichRep[duplicated(whichRep$inv), "inv"]
    
    subRep<-whichRep[!(whichRep$inv %in% heteroRep) & whichRep$ChrRegion %in% repDens,"inv" ]
    
    # Take info for selected rep density
    genoInfo<-genome[genome$ChrRegion %in%  repDens,]
    rownames(genoInfo)<-paste(genoInfo$Chromosome, genoInfo$Start, genoInfo$End, sep = "_")
    
    # Replace info for inv windows
    invInfo<-invs[invs$inv %in% as.character(subRep),]
    invInfo$tags<-paste(invInfo$Chromosome, invInfo$Start, invInfo$End, sep = "_")
    
    genoInfo[invInfo$tags, "ChrRegion" ]<-invInfo$inv
    
    ###MAK ELOOP
    allinvsamples<-data.frame()
    for (inversion in subRep){
      
      sampleSize<-nrow(invInfo[invInfo$inv == inversion,])
      
      # Make subset  
      toSam<-genoInfo[genoInfo$ChrRegion %in%  repDens,]
      toSam$previous<-toSam$Start - c(0, toSam$Start[1:length(toSam$Start)-1])
      
      # Make sure it is sorted
      toSam<-toSam[order(toSam$Chromosome, toSam$Start),]
      
      # Give it rownames
      rownames(toSam)<-c(1:nrow(toSam))
      
      # Select sampleable rownames
      nonums<-c()
      nostarters<-as.numeric(rownames(toSam[toSam$previous != 10000,]))-sampleSize+1
      for(n in nostarters){
        nonums<-c(nonums, c(n:(n+sampleSize-2)))
      }
      
      options<-as.numeric(rownames(toSam)[!(as.numeric(rownames(toSam)) %in% nonums)])
  
      # Make random samples
      samples<-data.frame()
      
      starters<-sample(options, nSamples, replace = F)
      samples<-data.frame()
        for(i in starters){
          sample<-toSam[seq(i,i+sampleSize-1),]
          sample$name<-i
          samples<-rbind(samples,sample)
        }
      samples$inv<-inversion
      allinvsamples<-rbind(allinvsamples, samples)
    }
    
return(list(allinvsamples,genoInfo[!(genoInfo$ChrRegion %in%  repDens),] ))
}
```


```{r}
# Test function
repDens<-"High005"
nSamples<-1000

samples<-sampleGenome(invs=invs, genome=genome, repDens=repDens, nSamples=nSamples)


```



I can also make a meta-analysis of:
min and max weighted recRate for inv vs the same for an equivalent number of windows randomly selected

OK BUT
what they comapre is the ratio between population and pedigree, so I need the same windows for each...
1.1 For each window make max, min, weighted RecRate (with fillWindowMeasurements), FOR THE OTHER WINDOW SET (I make both just in case one of them doesn't work)

```{bash}
TODAY="defTest"
DIR="analysis/${TODAY}_LocationPatterns/"

#CONDITIONS="avgBherer CEUSpence"

ENVPY="/home/rgomez/anaconda3/bin/python"

#Bherer
CONDITION="CEUSpence"
MAP="data/Bherer_Refined_genetic_map_b37_processed/sexavg_allChr.bed"
FILWINDOW="$DIR/slWins/slWins_${CONDITION}_filtered.bed"

$ENVPY code/python/fillWindowMeasurements.py -m $MAP -w $FILWINDOW -o $DIR/slWins/flipped -d $CONDITION


#Bherer
CONDITION="avgBherer"
MAP="data/SpenceSong_hg19_recMaps_processed/CEU_recombination_map_hg19_allChr.bed"
FILWINDOW="$DIR/slWins/slWins_${CONDITION}_filtered.bed"

$ENVPY code/python/fillWindowMeasurements.py -m $MAP -w $FILWINDOW -o $DIR/slWins/flipped -d $CONDITION


```

Now I'll use the avgBherer with Spence map which is the one that worked best:

```{r}
library(ggplot2)
condition<-"avgBherer"
repDens<-"High005"

# Upload genome measurements for a given condition
genome_ped<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/fillWindowMeasurements/",condition,"/windowData.txt"), header = T, stringsAsFactors = F)
genome_pop<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/flipped/fillWindowMeasurements/",condition,"/windowData.txt"), header = T, stringsAsFactors = F)

# Ratio of the two RATES ped/pop
genome<-merge(genome_ped, genome_pop, by = c("Chromosome", "Start", "End", "ChrRegion"), suffixes = c("_ped", "_pop") )
genome$WAvgRate_ped_pop<-genome$WAvgRate_ped/genome$WAvgRate_pop

# Upload inversion info
invs<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/slWins_",condition,"_invs.bed"), stringsAsFactors = F)
colnames(invs)<-c("Chromosome", "Start", "End", "ChrRegion", "order", "invChromosome", "invStart", "invEnd", "inv")

#Make samples
samples<-sampleGenome(invs=invs, genome=genome, repDens =repDens, nSamples = 1000)
sampleMeasurements<-samples[[1]]
invMeasurements<-samples[[2]]

mins_dist<-aggregate( WAvgRate_ped_pop ~ name+inv, sampleMeasurements, min)
maxs_dist<-aggregate( WAvgRate_ped_pop ~ name+inv, sampleMeasurements, max)
randist<-merge(mins_dist, maxs_dist, by = c("name", "inv"), suffixes = c(".min", ".max"))
randist<-melt(randist, id.vars = c("inv", "name"))

mins_ori<-aggregate(WAvgRate_ped_pop ~ ChrRegion, invMeasurements, min )
maxs_ori<-aggregate(WAvgRate_ped_pop ~ ChrRegion, invMeasurements, max )
oridist<-merge(mins_ori, maxs_ori, by = c( "ChrRegion"), suffixes = c(".min", ".max"))
oridist<-melt(oridist, id.vars = c("ChrRegion"))
colnames(oridist)<-c("inv", "variable", "value")

ggplot()+geom_boxplot(data=randist, aes(x = inv, y = value))+
  geom_point(data=oridist, aes(x = inv, y = value), color = "red")+
  scale_y_log10()+
  facet_grid(variable~.)

```


SHOULD REPEAT THOSE ANALYSES THAT DEPEND ON FILLWINDOWMEASUREMENTS!

# Invented method

## Step 1
I will need: sliding windows with min, max, weighted recRate (done)

```{r}
library(ggplot2)
condition<-"avgBherer"

# Upload genome measurements for a given condition
genome<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/fillWindowMeasurements/",condition,"/windowData.txt"), header = T, stringsAsFactors = F)

```

These sliding windows will be used to compare rec measurements inside the actual and simulated inversions.
## Step 2 Generate inv coordinate tables (actual and simulated)
OJO! 

Why do I sample the genome and not all the possible numbers? The only reason is because a vector with all the positions in the genome would be very large, while sampling windows is a good approximation. I exclude those windows overlapping with inversions and I leave space at the start and the end of the chromosome to not cut the inversion, but I do not leave space for flanking regions because it could happen in real life that I do not have enough space for flanking regions of a determinate size. 

> In addition, I want to make sure that permutations do not overlap other invesions, even if I'm looking only at big invs. This is somethong that Basatena does to avoid biases. 

```{r}

# These are INCLUSIVE made with bedtools
invs<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/slWins_",condition,"_invs.bed"), stringsAsFactors = F)
colnames(invs)<-c("Chromosome", "Start", "End", "ChrRegion", "order", "invChromosome", "invStart", "invEnd", "inv")

```

```{r}
# OJO! measurements overlap a lot for invs < 100 kb (sliding window size). They use originally this size for invs > 500kb
# I will make a """general""" dataset (> 100kb) and a "paper equivalent" dataset (> 500kb)
#This will make simulations go faster

biginvs<-unique(invs[invs$invSize >= 100000, "inv"])
verybiginvs<-unique(invs[invs$invSize >= 500000, "inv"])

```

```{r function_sampleGenome}
sampleGenome<-function(toSample, invInfo, genoInfo, repDens, nSamples){
  slidingSize<-10000
  ####################################################
    # toSample is the list of invs we want to simulate
    # invInfo is the list of all invs (we want to avoid overlaps with these)
    # genome is the dataset to sample from
    # repDens can be "same" to resample in the main repDens profile of the real inv or "all" to resample in all the genome
    # nSamples is sample size
  ####################################################
  
    # TAKE INVERSION SAMPLE TEMPLATES
      toSampleInfo<-invInfo[invInfo$inv %in% toSample,]
    
    # DETERMINE REPEAT STATUS for sample templates
      repeatsTable<-data.frame(table(toSampleInfo$inv, toSampleInfo$ChrRegion))
      repeatsTable<-merge(aggregate(Freq ~ Var1, data = repeatsTable, FUN = "max"), repeatsTable, all.x = T)
      
    # GENOME FILTERIGNG
      # Tag genome
      rownames(genoInfo)<-paste(genoInfo$Chromosome, genoInfo$Start, genoInfo$End, sep = "_")
     
      # Tag inv list
      invInfo$tags<-paste(invInfo$Chromosome, invInfo$Start, invInfo$End, sep = "_")
      
      # Replace Repeat info to avoid ALL inv regions
      genoInfo[invInfo$tags, "ChrRegion" ]<-invInfo$inv
      
      # Now, for each used repeat status OR for all the genome, we want to make a list of all the sampleable sites
      sampleables<-list()
      if(repDens == "same"){
        for (repD in unique(repeatsTable$Var2)){
         sampleables[[repD]]<-genoInfo[genoInfo$ChrRegion == repD,] 
        }
      }else{
        sampleables[["all"]]<-genoInfo[genoInfo$ChrRegion %in%  unique(repeatsTable$Var2) ,]
      }
    
    # MAKE LOOP
    allinvsamples<-data.frame()

    for (inversion in toSample){
      
      # Calculate number of windows
      sampleSize<-nrow(invInfo[invInfo$inv == inversion,])
      
      # Select subset and prepare to calculate excluded coords  
      whereToSample<-ifelse(repDens == "all", "all", as.character(repeatsTable[repeatsTable$Var1 == inversion,"Var2"]))
      toSam<-sampleables[[whereToSample]]     
      toSam$previous<-toSam$Start - c(0, toSam$Start[1:length(toSam$Start)-1]) #mark previous coord
      toSam<-toSam[order(toSam$Chromosome, toSam$Start),]#sort
      rownames(toSam)<-c(1:nrow(toSam)) # give rownames
      
      # Select sampleable rownames
      nonums<-c()
      nostarters<-as.numeric(rownames(toSam[toSam$previous != slidingSize,]))-sampleSize+1
      for(n in nostarters){
        nonums<-c(nonums, c(n:(n+sampleSize-2)))
      }
      
      options<-as.numeric(rownames(toSam)[!(as.numeric(rownames(toSam)) %in% nonums)])
  
      # Make random samples
      starters<-sample(options, nSamples, replace = T)
      
      # Generate a table like the invInfo table
      samples<-toSam[starters, c("Chromosome", "Start", "End", "ChrRegion")]
      samples$End<-samples$Start + unique(invInfo[invInfo$inv == inversion,"invSize"])
      samples$inv<-paste(inversion , c(1:nrow(samples)), sep = "_")
      
      
      # Add to table
      allinvsamples<-rbind(allinvsamples,samples)
    }
    colnames(allinvsamples)<-c("Chromosome", "invStart", "invEnd", "ChrRegion", "inv")
return(allinvsamples)
}
```

```{r}
#Bigincs
biginvs_permutations_all<-sampleGenome(biginvs, invs, genome, "all", 10000)
biginvs_permutations_same<-sampleGenome(biginvs, invs, genome, "same", 10000)

# Verybiginvs
verybiginvs_permutations_all<-sampleGenome(verybiginvs, invs, genome, "all", 10000)
verybiginvs_permutations_same<-sampleGenome(verybiginvs, invs, genome, "same", 10000)

```


## Step 3 Core method (take measurements of avg, min, max fold change)

Since I will need permutations to compare to, all the following method should be grouped in easy-use functions that take as input inversions coordinates (actual or simulated). 
 I will also need the measurements for the windows outside, in fixed and proportional modes. In Seich Basatena paper they use fixed 0.75Mb = 750kb for a 0.75Mb simulated inv. I could use something near the median inv length, maybe (not the average bc there are very big invs). 

```{r}
summary(invs$invEnd - invs$invStart )
```


So, 200kb it is. 
Since I do not leave space for flanking regions because it could happen in real life that I do not have enough space for flanking regions of a determinate size, the simulations with negative sides will be deleted. However, this only works with the start of the chromosome. Then, I need to put another filter for the overlaps... so I need an overlap measurement.

```{r}
library(reshape2)
#Make an adjacent list
makeAdjacents<-function(invs, store){
  #############################################
  # Invs is a table, and I need the columns invEnd, invStart, "Chromosome, inv
  #####################################
  invs$invSize<-invs$invEnd - invs$invStart
  FixSize<-200000
  
   ## Proportional
  invs$outPropL.Start<-invs$invStart - invs$invSize
  invs$outPropL.End<-invs$invStart
  invs$outPropR.Start<-invs$invEnd 
  invs$outPropR.End<-invs$invEnd + invs$invSize
   ## Fixed
  invs$outFixL.Start<-invs$invStart - FixSize
  invs$outFixL.End<-invs$invStart
  invs$outFixR.Start<-invs$invEnd 
  invs$outFixR.End<-invs$invEnd + FixSize
  
  # Adjacents
  outs<-melt(invs[, c("Chromosome", "inv", grep("out", colnames(invs), value = T))], id.vars = c("Chromosome", "inv"))
  outs<-cbind(outs, stringr::str_split_fixed(outs$variable, "\\.", 2))
  outs<-reshape(outs, direction = "wide", idvar = c("Chromosome" , "inv", "1"), timevar = "2")
  outs$name<-paste(outs$inv, outs$`1`, sep = "_")
  outs<-outs[,c("Chromosome", "value.Start", "value.End", "name")]
  
  # Remove negative adjacents
  outs<-outs[outs$value.Start >0,]
  
  write.table(outs, store, quote = F, sep = "\t", row.names = F, col.names = F)
  return(outs)
}
```

Calculate adjacent tables
```{r}
dir.create( "analysis/defTest_LocationPatterns/slWins/adjacentCoords/")

invs_sides<-makeAdjacents(unique(invs[,c("Chromosome", "invStart", "invEnd" , "inv")]), "analysis/defTest_LocationPatterns/slWins/adjacentCoords/invs.bed")

#Bigincs
biginvs_permutations_all_sides<-makeAdjacents(biginvs_permutations_all, "analysis/defTest_LocationPatterns/slWins/adjacentCoords/biginvs_permutations_all.bed")
biginvs_permutations_same_sides<-makeAdjacents(biginvs_permutations_same, "analysis/defTest_LocationPatterns/slWins/adjacentCoords/biginvs_permutations_same.bed")

# Verybiginvs
verybiginvs_permutations_all_sides<-makeAdjacents(verybiginvs_permutations_all, "analysis/defTest_LocationPatterns/slWins/adjacentCoords/verybiginvs_permutations_all.bed")
verybiginvs_permutations_same_sides<-makeAdjacents(verybiginvs_permutations_same, "analysis/defTest_LocationPatterns/slWins/adjacentCoords/verybiginvs_permutations_same.bed")

unique(biginvs_permutations_all[grep("HsInv0501", biginvs_permutations_all$inv), c("Chromosome", "invStart")])

```


Take windows ovelapping outside coords. I don't know if these simulations are valid for both maps!!I took the samples in the Bherer map, which has less regions avaiable, so it should be more or less universal. I require the overlap to be reported to delete those entries that go beyond the chromosome limits. The size for each condition was about 8Gb and I couldn't load it all together, so I changed the loops to have 1 overlap file per original bedfile. 

```{bash}
TODAY="defTest"
DIR="analysis/${TODAY}_LocationPatterns/"
CONDITIONS="avgBherer CEUSpence"
OUTDIR="${DIR}/slWins/"
INVCOORDS="${DIR}/slWins/adjacentCoords/"
#HEADER="Chromosome\tStart\tEnd\tChrRegion"
mkdir -p "$OUTDIR/adjacentMeasurements/"

for CONDITION in $CONDITIONS; do
  for INVCOORD in $(ls $INVCOORDS); do

   WINDOW="$OUTDIR/slWins_${CONDITION}.bed"
   OVERLAPS="$OUTDIR/adjacentMeasurements/slWaC_${CONDITION}_${INVCOORD}"
  
   bedtools intersect -wo -a $WINDOW -b $INVCOORDS/$INVCOORD   > $OVERLAPS
  done
done
```

Now that I have everything in a single table, I can make all the comparisons, but I want to make sure that it can be repeated for another dataset (each file)

```{r}
# As if this were a function
#######################
# Here the inputs

# The file shuold be processed inside the function because they are large files
file<-"analysis/defTest_LocationPatterns/slWins/adjacentMeasurements/slWaC_avgBherer_verybiginvs_permutations_all.bed"
condition<-"avgBherer"

######################

# These are INCLUSIVE made with bedtools
out<-read.table(file, stringsAsFactors = F)
colnames(out)<-c("Chromosome", "Start", "End", "ChrRegion",  "outChromosome", "outStart", "outEnd", "outID", "overlap")

# Separate by types
out$type<-"Prop"
out[grep("Fix", out$outID), "type"]<-"Fix"

# Separate the inv name
out<-cbind(out, stringr::str_split_fixed(out$outID, "_out", 2))
colnames(out)[colnames(out) %in% c("1", "2")]<-c("inv", "place")

# Save lefts an rights
out$place<-gsub("Fix", "", gsub("Prop", "",  out$place))

# ADD MEASUREMENTS
# Upload genome measurements for a given condition
genome<-read.table(paste0("analysis/defTest_LocationPatterns/slWins/fillWindowMeasurements/",condition,"/windowData.txt"), header = T, stringsAsFactors = F)

# Merge by coordinate
out<-merge(out, genome)
rm(genome)

outPropMeasurements<-merge(out[out$type == "Prop",], genome)
avg<-aggregate(WAvgRate ~ inv, outPropMeasurements, "mean")
min<-aggregate(WAvgRate ~ inv, outPropMeasurements, "min")
max<-aggregate(WAvgRate ~ inv, outPropMeasurements, "max")

outPropMesSum<-merge(merge(max, min, by = "inv", suffixes = c("_max", "_min")), avg, by = "inv")

# COMBINED OUT FIXED MEASUREMENTS
outFixMeasurements<-merge(out[out$type == "Fix",], genome)
avg<-aggregate(WAvgRate ~ inv, outFixMeasurements, "mean")
min<-aggregate(WAvgRate ~ inv, outFixMeasurements, "min")
max<-aggregate(WAvgRate ~ inv, outFixMeasurements, "max")

outFixMesSum<-merge(merge(max, min, by = "inv", suffixes = c("_max", "_min")), avg, by = "inv")


```



## Step 4 Apply method running function, make summaries

## --- this following goes to Step 3 mainly in the form of functions ---



```{r}
# IN MEASUREMENTS

inMeasurements<-merge(invs, genome )
avg<-aggregate(WAvgRate ~ inv, inMeasurements, "mean")
min<-aggregate(WAvgRate ~ inv, inMeasurements, "min")
max<-aggregate(WAvgRate ~ inv, inMeasurements, "max")

inMesSum<-merge(merge(max, min, by = "inv", suffixes = c("_max", "_min")), avg, by = "inv")

# COMBINED OUT PROPORTIONAL MEASUREMENTS

outPropMeasurements<-merge(out[out$type == "Prop",], genome)
avg<-aggregate(WAvgRate ~ inv, outPropMeasurements, "mean")
min<-aggregate(WAvgRate ~ inv, outPropMeasurements, "min")
max<-aggregate(WAvgRate ~ inv, outPropMeasurements, "max")

outPropMesSum<-merge(merge(max, min, by = "inv", suffixes = c("_max", "_min")), avg, by = "inv")

# COMBINED OUT FIXED MEASUREMENTS
outFixMeasurements<-merge(out[out$type == "Fix",], genome)
avg<-aggregate(WAvgRate ~ inv, outFixMeasurements, "mean")
min<-aggregate(WAvgRate ~ inv, outFixMeasurements, "min")
max<-aggregate(WAvgRate ~ inv, outFixMeasurements, "max")

outFixMesSum<-merge(merge(max, min, by = "inv", suffixes = c("_max", "_min")), avg, by = "inv")


```

And now the fold change for the measurments

```{r}
foldChanges<-merge(merge(outFixMesSum, outPropMesSum, by = "inv", suffixes = c("_outFix", "_outProp")), inMesSum, by = "inv")

foldChanges$FC_min_outFix<-foldChanges$WAvgRate_min/foldChanges$WAvgRate_min_outFix
foldChanges$FC_max_outFix<-foldChanges$WAvgRate_max/foldChanges$WAvgRate_max_outFix
foldChanges$FC_avg_outFix<-foldChanges$WAvgRate/foldChanges$WAvgRate_outFix
foldChanges$FC_min_outProp<-foldChanges$WAvgRate_min/foldChanges$WAvgRate_min_outProp
foldChanges$FC_max_outProp<-foldChanges$WAvgRate_max/foldChanges$WAvgRate_max_outProp
foldChanges$FC_avg_outProp<-foldChanges$WAvgRate/foldChanges$WAvgRate_outProp


```


```{r}
# OJO! measurements overlap a lot for invs < 100 kb (sliding window size). They use originally this size for invs > 500kb
# I will make a """general""" dataset (> 100kb) and a "paper equivalent" dataset (> 500kb)

biginvs<-unique(invs[invs$invSize >= 100000, "inv"])
verybiginvs<-unique(invs[invs$invSize >= 500000, "inv"])

```

```{r}
biginvs_fc<-foldChanges[foldChanges$inv %in% biginvs,]

verybiginvs_fc<-foldChanges[foldChanges$inv %in% verybiginvs,]

```


