
## Summary

This test is to check whether the spence map is biased as LdHat, as Bherer map, or not biased at all.
There are two analyses:
* Map vs. map: I compare ped/pop ratios (maximum and minimum) with expectations to make a p-value, and check how extreme are ratios compared to expected for each inversion.
* In vs. out: I compare (average, maximum and minimum) in/out ratios with expectations to make p-value, and check how the ratio p-value is affected by inversion frequency on each map. 

I have preliminary results for both map vs. map and in vs. out made in a sliding window map, but I also want to compare in vs. out with raw data. Sliding window map will be calculated probabilistically (average of rates weighted by map window size). Inversions containing >80% of a window map will be included in the map vs. map analysis. For the in vs. out analysis, windows at breakpoints can be deleted, also option 2: unless they overlap with >80% of either option, in which case they will be assigned to only in or only out. 

## Analysis

### Setup the directory

```{r}
load("analysisparams.Rdata")
outDir<-paste0("analysis/", today, "_LocationPatterns/04_MapValidityTest/")
dir.create(outDir)
```

### Make sliding window maps

STEP1 
Make sliding window map coordinates. We want to use
- strictly common workspace
- add centromere confidence interval at least for Spence

This script makes a workspace for Spence and a workspace for Bherer.
As for 20 oct 2022, the divideChromosomes script adds 15kb at either side of the centromere (30kb).
after making a workspace for each map, we want to make a consensus map with strictly the overlapping region.

STEP2
Divide workspace into sliding windows (100kb sliding windows at 10kb intervals). In previous versions of this analysis we took into account repeat regions, but we will keep it simple here. 

STEP3
Fill windows with information. As for 20 oct 2022, fillWindowMeasurements_avgonly.py takes all overlapping coordinates, cuts overhanging coordinates to fit window, calculates average crossover rate weighing by the proportion each coordinate occupies in the window. 

```{bash}

TODAY="defTest"
ENVPY="/home/rgomez/anaconda3/bin/python"
OUTDIR="analysis/${TODAY}_LocationPatterns/04_MapValidityTest"

#### STEP 1 ####
# MAKE COMMON WORKSPACE + CENTROMERE CI

parallel $ENVPY code/python/divideChromosomes.py -m {1} -b data/cytoBand.txt -f 1 -s 0 -r 0 -o $OUTDIR -d {2} ::: data/Bherer_Refined_genetic_map_b37_processed/sexavg_allChr.bed  data/SpenceSong_hg19_recMaps_processed/CEU_recombination_map_hg19_allChr.bed :::+ avgBherer CEUSpence

WORKDIR="${OUTDIR}/workspace_windows"
mkdir -p $WORKDIR

parallel grep -v '"cen"' "$OUTDIR/divideChromosomes/{}_fixedArms_1/workspace.txt" '|' awk '-v OFS="\t"' \''NR>1{print $4,$1,$2}'\' '>'  "$WORKDIR/workspace_{}.bed" ::: CEUSpence avgBherer

# Now I want to make a consensus workspace that takes only the common region.
bedtools intersect -a "$WORKDIR/workspace_CEUSpence.bed" -b "$WORKDIR/workspace_avgBherer.bed" > "$WORKDIR/workspace_common.bed"

#### STEP 2 ####
# MAKE SLIDING WINDOWS FOR COMMON WORKSPACE
bedtools makewindows -b "${WORKDIR}/workspace_common.bed" -w 100000 -s 10000  > "${WORKDIR}/slidingWindows.bed"
echo -e "Chromosome\tStart\tEnd\tChrRegion" > "${WORKDIR}/slidingWindows_header.bed"
cat "${WORKDIR}/slidingWindows.bed" >> "${WORKDIR}/slidingWindows_header.bed"

#### STEP 3 ####
MAPDIR="${OUTDIR}/recombinationMap_sliding"
mkdir -p $MAPDIR

parallel $ENVPY code/python/fillWindowMeasurements_avgonly.py -m {1} -w  "${WORKDIR}/slidingWindows_header.bed" -o ${MAPDIR} -d {2} ::: data/Bherer_Refined_genetic_map_b37_processed/sexavg_allChr.bed data/SpenceSong_hg19_recMaps_processed/CEU_recombination_map_hg19_allChr.bed :::+ Bherermap_sliding Spencemap_sliding


```

### Prepare input - inversions

```{r}

# Add libraries
pacman::p_load(vcfR, ggplot2)

# Setup 
dir.create(paste0(outDir,"/inversionList"), showWarnings = F)

# Read  VCF info
vcf<-read.vcfR("data/InversionGenotypes_processed/allgenotypes.vcf")
vcfInfo<-vcfR2tidy(vcf, info_only = TRUE)$fix

# Remove Del genotypes
vcfInfo<-data.frame(vcfInfo)

# Reformulate chromosomes
vcfInfo<-vcfInfo[!grepl("Del", vcfInfo$ID),]
vcfInfo$Chrom<-paste0("chr", vcfInfo$CHROM)

# Make bedfile
write.table(vcfInfo[,c("Chrom", "POS", "END", "ID")], paste0(outDir,"inversionList/133invs.bed"), quote = F, sep = "\t", row.names = F, col.names = F)

# Make individual bedifles as well
for (i in rownames(vcfInfo)){
  write.table(data.frame(vcfInfo[i,c("Chrom", "POS", "END", "ID")]), paste0(outDir,"inversionList/",vcfInfo[i,"ID"],".bed"), quote = F, sep = "\t", row.names = F, col.names = F)
}
```

### Calculate ratios and expectations


```{bash}

# This is for sliding window maps, requires 80% of overlap with at least one window for the inversion to be analyzed

mkdir -p "analysis/defTest_LocationPatterns/04_MapValidityTest/ratioExpectations"

parallel  Rscript code/rscript/checkMapValidity_sliding.R {} 200000 1000 0.8 "analysis/defTest_LocationPatterns/04_MapValidityTest/recombinationMap_sliding/fillWindowMeasurements/Bherermap_sliding/windowData.txt" "analysis/defTest_LocationPatterns/04_MapValidityTest/recombinationMap_sliding/fillWindowMeasurements/Spencemap_sliding/windowData.txt" "analysis/defTest_LocationPatterns/04_MapValidityTest/ratioExpectations/{/.}.txt" ::: analysis/defTest_LocationPatterns/04_MapValidityTest/inversionList/HsInv* > "analysis/defTest_LocationPatterns/04_MapValidityTest/ratioExpectations/unanalysed.txt"


```

```{bash}

# This is for sliding window maps, requires 80% of overlap with at least one window for the inversion to be analyzed

mkdir -p "analysis/defTest_LocationPatterns/04_MapValidityTest/ratioExpectations_strict"

parallel  Rscript code/rscript/checkMapValidity_sliding.R {} 200000 1000 1 "analysis/defTest_LocationPatterns/04_MapValidityTest/recombinationMap_sliding/fillWindowMeasurements/Bherermap_sliding/windowData.txt" "analysis/defTest_LocationPatterns/04_MapValidityTest/recombinationMap_sliding/fillWindowMeasurements/Spencemap_sliding/windowData.txt" "analysis/defTest_LocationPatterns/04_MapValidityTest/ratioExpectations_strict/{/.}.txt" ::: analysis/defTest_LocationPatterns/04_MapValidityTest/inversionList/HsInv* > "analysis/defTest_LocationPatterns/04_MapValidityTest/ratioExpectations_strict/unanalysed.txt"


```

```{r upload_files}
dir<-"ratioExpectations_strict"
files<-grep("HsInv",list.files(paste0(outDir, dir, "/")), value = T)

table<-data.frame()
for (i in files){
   table<-rbind(table, read.table(paste0(outDir,dir,  "/", i), header = T))
}

baseCols<-c("inv", "Chromosome", "Start", "End", "CIStart", "CIEnd")

mapmap<-table[,c(baseCols, grep("PedPop", colnames(table), value=T))]
mapmap<-unique(mapmap)
inout<-table[,c( grep("PedPop", colnames(table), value=T, invert = T) )]
```



### Map vs. map
I compare ped/pop ratios (average, maximum and minimum) with expectations to make a p-value, and check how extreme are ratios compared to expected for each inversion.
Sliding windows
Inversions containing >80% of a window map

```{r}

binomPval<-function( column, threshold, type){
  samples<-length(column)
  
  # If a max has a p-value >= 0.95, it has at must 5% of the distribution > observed; it is within the 5% higher in the dist
  #If a min has a p-value <= 0.05, it has at must 5% of the distribution < observed; it is within the 5% lower in the dist
  success<-ifelse(type =="max", sum(column >= 1-threshold), sum(column <= threshold)) 
  
  pval<-pbinom(success-1, size = samples, prob = threshold, lower.tail = F) #Probabilidad de n o mas successes (= probabilidad de mas de n-1)
  
  pval<-signif(pval,4)

  return(list(s = success, p = pval))
  }
    
# Make table as in the original paper (~al-Basatena2017)
mapmap["Inversion length (Mb)"]<-round((mapmap$End-mapmap$Start)/1000000, 2)
mapmap["% of empirical dist. < min ped/pop ratio"]<-mapmap$minPedPop_inside_pval*100
mapmap["% of empirical dist. > max ped/pop ratio"]<-(1-mapmap$maxPedPop_inside_pval)*100

print(mapmap[order(mapmap$`Inversion length (Mb)`, decreasing = T),c("inv", "Inversion length (Mb)", "% of empirical dist. > max ped/pop ratio", "% of empirical dist. < min ped/pop ratio")])

# Make p-values for the columns - hace lo mismo
# binomPval(mapmap$`% of empirical dist. < min ped/pop ratio`, 0.05 ,"min")
# binomPval(mapmap$`% of empirical dist. < min ped/pop ratio`, 0.01 ,"min")
# binomPval(mapmap$`% of empirical dist. > max ped/pop ratio`, 0.05, "min") #porque esta transformado
# binomPval(mapmap$`% of empirical dist. > max ped/pop ratio`, 0.01, "min") #porque esta transformado

# Test with original columns
min05<- binomPval(mapmap$minPedPop_inside_pval, 0.05 ,"min")
min01<- binomPval(mapmap$minPedPop_inside_pval, 0.01 ,"min")
max05<- binomPval(mapmap$maxPedPop_inside_pval, 0.05, "max")
max01<- binomPval(mapmap$maxPedPop_inside_pval, 0.01, "max")

totalinvs<-nrow(mapmap)

sprintf("The enrichment of values in the top 5%% of the empirical minimum ped/pop distribution among inversions has P = %g (%i of %i), and that in the top 1%% has P = %g (%i of %i), assuming binomial distribution.", min05$p, min05$s, totalinvs, min01$p, min01$s, totalinvs)

sprintf("The enrichment of values in the top 5%% of the empirical maximum ped/pop distribution among inversions has P = %g (%i of %i), and that in the top 1%% has P = %g (%i of %i), assuming binomial distribution.", max05$p, max05$s, totalinvs, max01$p, max01$s, totalinvs)

```



### In vs. out
I compare (average, maximum and minimum) in/out ratios with expectations to make p-value, and check how the ratio p-value is affected by inversion frequency on each map. 
Sliding windows and raw data
option 1: windows at breakpoints can be deleted, also 
option 2: unless they overlap with >80% of either option, in which case they will be assigned to only in or only out

```{r}

library(vcfR)
vcf<-read.vcfR("data/InversionGenotypes_processed/allgenotypes.vcf")
vcfInfo<-vcfR2tidy(vcf, info_only = TRUE)$fix

freqs<-vcfInfo[,c("ID", "AF_EUR", "AF")]
freqs<-merge( freqs,inout, by.x = "ID", by.y = "inv")

library(ggpubr)
library(gridExtra)


ggplot(freqs, aes(x = as.numeric(AF_EUR), y = Fold_change_inout_pval, color=map, fill = map))+geom_point()+geom_smooth(method = "lm", alpha = 0.1 )+stat_cor(method = "pearson", label.x = 0.4)+ggtitle("Influence of frequency in CO rate reduction")+labs(x = "Frequency in europeans", y = "p-value of min rec.rate inside/outside ratio")+facet_grid(variable~.)+ylim(0,1)


```

