<!--
---
 title: "Ordinal logistic model on large windows data - chapter 2"
 author: "Ruth Gómez Graciani"
 output:
   pdf_document: default
 header-includes:
  \usepackage{float}
  \floatplacement{figure}{H}
---
--> 
```{r, include=FALSE}
if (!require("pacman")) install.packages("pacman")
# set conditions for code output
knitr::opts_chunk$set(message=FALSE, warning=FALSE, error=FALSE,fig.pos = "!H")
showcode=FALSE #for include or echo - echo=showcode 
runbash=TRUE # for eval=runbash
runR=TRUE #for eval=runR
datadir="data/fillWindowMeasurements/chromRegions/"
```

Raw data: 

```{r c_loadData, echo=showcode, eval=runR}

pacman::p_load(foreign, ggplot2, MASS, Hmisc, reshape2, ordinal)
winRegions <- read.table(paste0(datadir, "windowData.txt"), header = T)

# Make factors for inversions
winRegions$invCenters<-factor(winRegions$invCenters, ordered = T)
winRegions$NHCenters<-factor(winRegions$NHCenters, ordered = T)
winRegions$NAHRCenters<-factor(winRegions$NAHRCenters, ordered = T)

# Change dimension of variables
winRegions$Length.Mb <- winRegions$Length.bp./1000000
winRegions$WAvgRate.perMb<- winRegions$WAvgRate * 1000000
winRegions$log10RepCounts<-log10(winRegions$allRepCounts)
winRegions<-winRegions[,c("Chromosome", "Start", "End", "Color", "invCenters", "NHCenters", "NAHRCenters", "Length.Mb", "allRepCounts", "log10RepCounts", "WAvgRate.perMb" )]

winRegions$ChromType<-factor(ifelse(winRegions$Chromosome == "chrX", "X", "A"))

# Show table
knitr::kable(head(winRegions))
```

For each window, I calculated the number of total inversions, NH inversions, and NAHR inversions, the window length in Mb, number of repeats and the average recombination rate in cM/Mb.

I want to perform Ordinal Logistic Regressions on different subsets of the data. The assumptions of the Ordinal Logistic Regression are as follow:

1. The dependent variable is ordered.
1. One or more of the independent variables are either continuous, categorical or ordinal.
1. No multi-collinearity.
1. Proportional odds.

I show the data distributions in the figure below. The inversion counts have only a number of possible options, so they can be considered an ordinal variable. The independent variables are continuous and categorical, so assumptions 1 and 2 are satisfied

```{r c_varDistribution, eval=runR, echo=showcode,  fig.cap="Distribution of variables.",  fig.height=5.5 }

pacman::p_load(ggdist, ggplot2, gghalves, reshape2, patchwork)
  
windowData <- winRegions
  # Put chromosome factor in order
  windowData$Chromosome <- factor(windowData$Chromosome, levels = paste(rep("chr", 22), as.character(c(1:22)), sep = ""))
  # windowData$chromID <- windowData$winID <- NULL
  
  # Melt data
  windowDataMelted <- melt(windowData, id.vars = c( "Chromosome", "Start", "End", "Color", "ChromType")) 
  windowDataMelted$value <- as.numeric(windowDataMelted$value)
  
  # Group info
  windowDataMelted$dataGroup<- ifelse(windowDataMelted$variable %in% c("invCenters", "NHCenters", "NAHRCenters"), "Inversions", 
                                      ifelse(windowDataMelted$variable %in% c("allRepCounts", "intraRepCounts", "log10RepCounts"), "Repeats",
                                             ifelse(windowDataMelted$variable %in% c("WAvgRate"), "Weighted average recRate",
                                             ifelse(windowDataMelted$variable %in% c("WAvgRate.perMb"), "Weighted average cM/Mb",
                                             ifelse(windowDataMelted$variable %in% c("Length.Mb"), "Window length in Mb",
                                                    ifelse(windowDataMelted$variable %in% c("maxRate"), "Maximum recRate",  "Window length"))))))

  # Make list of plots
  plot_list<-list()
  for (group in unique(windowDataMelted$dataGroup)) {
    
    plotTable <- windowDataMelted[(windowDataMelted$value != -Inf) & (windowDataMelted$dataGroup == group),]
    plot_list[[group]] <- ggplot(plotTable, aes(x = variable, y = value))+
      # Half violin
      ggdist::stat_halfeye(adjust = .5, width = .6, .width = 0, justification = -.2, point_colour = NA) +
      # Boxplot 
      geom_boxplot(width = .1, outlier.shape = NA) +
      # Points
      gghalves::geom_half_point_panel(side = "l", range_scale = .6,  alpha = .5, aes(color = Color))+
      # scale_color_manual(values = c(rep("#3c7ae7",11),rep("#89b23e",11) ))+
      # Adjust coordinates
      coord_flip()+
      # coord_flip( xlim = c(1.3, NA))+
      # Adjust labels
      theme(axis.title.y = element_blank(), legend.position = "top")+
      # Title
      ggtitle(group)
  }
  name = "Distribution of variables"
  # Plot list of plots
  wrap_plots(plot_list)+ plot_annotation( title = name)+plot_layout(guides = 'collect') & theme(legend.position = 'bottom')


```

We see that some categories have low number of cases, so I will make a "3 or more" category when relevant.

```{r c_reassignCategories, echo=showcode, eval=runR}


alltab<-data.frame("CountGroups" = c(0))
for (category in c( "invCenters","NHCenters", "NAHRCenters")) {
  df<-data.frame(table(winRegions[,category]))
  colnames(df)<-c("CountGroups", category)
  alltab<-merge( alltab, df, by = "CountGroups", all=T)
}
knitr::kable(alltab, caption = "Original counts")


alltab<-data.frame("CountGroups" = c("Absence"))
for (category in c( "inv","NH", "NAHR")) {
  # Store vector
  fill<-winRegions[,paste0(category,"Centers")]
  # Reassign category
  category<- paste0(category,"Category")
  # Copy vector
  winRegions[,category]<-fill
  # Reassign levels
  lev<-levels(winRegions[,category])
  catcon<-length(lev)-2
  levels(winRegions[,category])<-  c(c("Absence", "Presence", "Presence"), rep("Abundance",catcon))
  df<-data.frame(table(winRegions[,category]))
  colnames(df)<-c("CountGroups", category)
  alltab<-merge( alltab, df, by = "CountGroups", all=T)
}
knitr::kable(alltab[order(alltab$CountGroups),], caption = "New counts")




```
\newpage
With these groups, I visualize the relationships between dependent and independent variables.

```{r c_groupDifferences, eval=runR, echo=showcode,  fig.cap="Potential effect of independent variables on the different types of invesions.",  fig.height=8 }

pacman::p_load(ggplot2, reshape2,  grid, gridExtra)

winRegions_long<-melt(winRegions, id.vars = c("Start", "End", "Chromosome", "invCenters", "NHCenters", "NAHRCenters", "invCategory", "NHCategory", "NAHRCategory","Color"))

option <- "append" # empty o append
# Make list of plots
  plot_list<-list()
  for (group in c("invCategory", "NHCategory", "NAHRCategory")) {
    
    # Make new table
    winRegions_long$plotgroup<-winRegions_long[,group]
    # winRegions_long$value <- as.numeric(winRegions_long$value)
    winRegions$plotgroup <- winRegions[,group]
    # Make main plot
    p <- ggplot(winRegions_long)+
      geom_boxplot(aes(x = plotgroup, group = plotgroup, fill = plotgroup, y = value ), outlier.alpha = 0 , alpha = 0.8)+
      geom_point(aes(x = plotgroup, y = value ), alpha = 0.3, position = "jitter")+
      facet_wrap(variable~., scales = "free")+
      xlab("")+ylab("Variable Value")+
      theme(legend.position = "none")+
      ggtitle(group)

  
    # Make secondary plot
    h<- ggplot(winRegions)+
      geom_bar(aes(x= plotgroup, fill = plotgroup), position ="dodge")+
      facet_grid(Color~. )+
      xlab("")+
      theme(legend.position = "none", axis.title.y = element_blank())+
      ggtitle("")
   

   
     if (option == "empty") {
           gp <- ggplotGrob(p)
          h_grob<-ggplotGrob(h)
  
          # Calculate empty area from main plot
          empty.area <- gtable_filter(gp, "panel", trim = F)
          empty.area <- empty.area$layout[sapply(empty.area$grob, function(x){class(x)[[1]]=="zeroGrob"}),]
          
          empty.area$t <- empty.area$t - 1 #extend up by 1 cell to cover facet header
          empty.area$b <- empty.area$b + 1 #extend down by 1 cell to cover x-axis

          # Add grob to main plot
          gp0 <- gtable_add_grob(x = gp,
                                 grobs = h_grob,
                                 t = min(empty.area$t),
                                 l = min(empty.area$l), 
                                 b = max(empty.area$b), 
                                 r = max(empty.area$r))

            # Print plot
            
            plot_list[[group]] <- arrangeGrob(gp0)
            
     } else if (option == "append") {
       
        plot_list[[group]] <- arrangeGrob(p,h, nrow = 1, widths = c(3, 1) )
        
     }

    }

  
  # Plot list of plots
  wrap_plots(plot_list, ncol=1)+ plot_annotation( title = "Differences in each chromosomal variable between inversion count groups")


```

\newpage
Finally, I will test assumption number 3, no multi-collinearity between independent variables. 

```{r c_correlations, eval=runR, echo=showcode,  fig.cap="Correlations between variables.",  fig.height=8.5 }
pacman::p_load(PerformanceAnalytics, GGally, cowplot)

# Select columns
cols<- c("Length.Mb", "allRepCounts", "WAvgRate.perMb", "Color")

# Setup corrplot
  diagonal <- function(data, mapping, ...){
    p<-ggplot(data=data, mapping = mapping)+
      geom_histogram(aes(y= ..density..), bins = 9, fill = "lightgray", color = "gray")+
      geom_density()
  }

 lines <- function(data, mapping, ...){
    p<-ggplot(data=data, mapping = mapping)+
      geom_smooth(method=lm, alpha=0)+
      geom_point(size=0.5)
 }
 
 p1<- ggpairs(winRegions,  mapping = aes(color = Color, alpha = 0.5),columns = cols, 
              # lower = list(continuous = wrap("smooth", size = 0.5)),
              lower = list(continuous = lines),
              diag = list(continuous = diagonal), 
              upper = list(continuous = wrap("cor", method = "pearson", size = 3)))+
   ggtitle("Pearson correlation")+theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   scale_fill_manual(values=c("#00AFBB", "#E7B800", "#FC4E07")) +scale_color_manual(values=c("#00AFBB", "#E7B800", "#FC4E07"))  
 
  p2<- ggpairs(winRegions,  mapping = aes(color = Color, alpha = 0.5),columns = cols, 
              # lower = list(continuous = wrap("smooth", size = 0.5)),
              lower = list(continuous = lines),
              diag = list(continuous = diagonal), 
              upper = list(continuous = wrap("cor", method = "spearman", size = 3)))+
   ggtitle("Spearman correlation")+theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   scale_fill_manual(values=c("#00AFBB", "#E7B800", "#FC4E07")) +scale_color_manual(values=c("#00AFBB", "#E7B800", "#FC4E07"))  

  # Print corrplots
 plot_grid(
   ggmatrix_gtable(p1),
   ggmatrix_gtable(p2),
   nrow = 2)
 
 
 
```

We see that our three variables are significantly correlated, but this does not confirm multi-collinearity. I perform a variance inflation factor test on the corresponging linear model to further check the multi-collinearity.

```{r c_multicoll, eval=runR, echo=showcode}
pacman::p_load("car")
# check VIF
if ("chrX" %in% winRegions$ChromType){
  modlm <- lm(as.numeric(as.character(invCenters)) ~ Length.Mb + allRepCounts + Color + WAvgRate.perMb+ChromType , data = winRegions)
knitr::kable(vif(modlm))
}else{
  modlm <- lm(as.numeric(as.character(invCenters)) ~ Length.Mb + allRepCounts + Color + WAvgRate.perMb , data = winRegions)
knitr::kable(vif(modlm))

}


```
The general rule of thumbs for VIF test is that if the VIF value is greater than 10, then there is multi-collinearity, so we can say that the third assumption (no multi-collinearity) is satisfied.

The proportional odds assumption will be tested for each model that we fit in the following analyses.

## Variable scalation (optional)

Standardized coefficients are useful in our case to compare effects of predictors reported in different units. The most straightforward way is using the Agresti method of standardization, applied with the `scale()` function. 


```{r c_scaleVars,  echo=showcode, eval=runR}
winRegions$Length.Mb.Scaled <- scale(winRegions$Length.Mb)[,1]
winRegions$allRepCounts.Scaled<-scale(winRegions$allRepCounts)[,1]
winRegions$WAvgRate.perMb.Scaled<-scale(winRegions$WAvgRate.perMb)[,1]


knitr::kable(as.data.frame(apply(winRegions[,c("Length.Mb", "Length.Mb.Scaled", "allRepCounts", "allRepCounts.Scaled", "WAvgRate.perMb", "WAvgRate.perMb.Scaled")],2,summary)))
```
Once the model is fitted, we can use the sd to transform scaled coefficients to natural coefficients and viceversa. 


```{r c_descaleVars, echo = showcode, eval=showcode}
# The two model options
fit.t1<-polr(invCenters~allRepCounts.Scaled, winRegions)
fit.t2<-polr(invCenters~allRepCounts, winRegions)

# We store the scaled coefficient
scaledcoef_rep<-coef(fit.t1)["allRepCounts.Scaled"]

# We store the SD
sd_rep <- sd(winRegions$allRepCounts)

# alculated natural coefficient = Scaled coefficient / sd
(naturalcoef_rep <- scaledcoef_rep / sd_rep)
# Natural coefficient
(calculated_naturalcoef_rep<-coef(fit.t2)["allRepCounts"])
```

<!-- We now store the data as an R object for the next document -->

```{r, echo=FALSE, eval=runR}
save(winRegions, file= "data/winRegions.RData")
```

