---
title: "Ordinal logistic model on large, classified windows data"
author: "Ruth Gómez Graciani"
output:
  pdf_document: default
header-includes:
 \usepackage{float}
 \floatplacement{figure}{H}
---

\tableofcontents

# Prepare the data

```{r child = '01PrepareTheData_femBherer.Rmd'}
```

\newpage


# Numerical categories

## Descriptive statistics

```{r child = '02ExploreTheData_numCategories.Rmd'}
```

\newpage

## Not scaled variables
### Total inversions model

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb + ChromType" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4, 0), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High", NA), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r child = '03invAnalysis.Rmd'}
```

\newpage


### NH inversions model

```{r, include=FALSE}
# Inversion type
yVar<-"NHCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb + ChromType" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4, 0), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High", NA), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r child = '03invAnalysis.Rmd'}
```

\newpage

### NAHR inversions model

```{r, include=FALSE}
# Inversion type
yVar<-"NAHRCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb + ChromType" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4, 0), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High", NA), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```

```{r child = '03invAnalysis.Rmd'}
```




## Scaled variables
### Total inversions model

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb.Scaled + allRepCounts.Scaled + Color + WAvgRate.perMb.Scaled + ChromType" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4,0), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High", NA), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r child = '03invAnalysis.Rmd'}
```

\newpage


### NH inversions model

```{r, include=FALSE}
# Inversion type
yVar<-"NHCategory"
# Scaled/notScaled
xVars<- "Length.Mb.Scaled + allRepCounts.Scaled + Color + WAvgRate.perMb.Scaled + ChromType" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4,0), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High", NA), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r child = '03invAnalysis.Rmd'}
```

\newpage

### NAHR inversions model

```{r, include=FALSE}
# Inversion type
yVar<-"NAHRCategory"
# Scaled/notScaled
xVars<- "Length.Mb.Scaled + allRepCounts.Scaled + Color + WAvgRate.perMb.Scaled + ChromType" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4,0), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High", NA), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```

```{r child = '03invAnalysis.Rmd'}
```



# Descriptive categories

## Descriptive statistics

```{r child = '02ExploreTheData_charCategories.Rmd'}
```

\newpage
## Not scaled variables
### Total inversions model

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb + ChromType" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4, 0), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High", NA), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r child = '03invAnalysis.Rmd'}
```

\newpage


### NH inversions model

```{r, include=FALSE}
# Inversion type
yVar<-"NHCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb + ChromType" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4, 0), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High", NA), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r child = '03invAnalysis.Rmd'}
```


\newpage

### NAHR inversions model

<!-- This cannot be done with ordinal logistic regression because we have only 2 categories, we would make a binomial logistic regression. -->

```{r, include=FALSE}
# Inversion type
yVar<-"NAHRCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb + ChromType"

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4,0), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High", NA), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```

```{r child = '03invAnalysis.Rmd'}
```


## Scaled variables
### Total inversions model

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb.Scaled + allRepCounts.Scaled + Color + WAvgRate.perMb.Scaled + ChromType" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4, 0), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High", NA), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r child = '03invAnalysis.Rmd'}
```

\newpage


### NH inversions model

```{r, include=FALSE}
# Inversion type
yVar<-"NHCategory"

# Scaled/notScaled
xVars<- "Length.Mb.Scaled + allRepCounts.Scaled + Color + WAvgRate.perMb.Scaled + ChromType" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4, 0), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High", NA), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r child = '03invAnalysis.Rmd'}
```


\newpage

### NAHR inversions model

This cannot be done with ordinal logistic regression because we have only 2 categories, we would make a binomial logistic regression.

```{r, include=FALSE}
# Inversion type
yVar<-"NAHRCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb + ChromType"

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4,0), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High", NA), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```

```{r child = '03invAnalysis.Rmd'}
```
