---
title: "Detection of large scale inversion location patterns with ordinal logistic regression"
author: "Ruth Gómez Graciani"
output:
  bookdown::pdf_document2: default
header-includes:
 \usepackage{float}
 \floatplacement{figure}{H}
 \floatplacement{table}{H}
---

\tableofcontents

\newpage
# Windows generation

```{r child = '01PrepareTheData_CEUSpence.Rmd'}
```

\newpage


# Data preparation

```{r child = '02ExploreTheData_numCategories.Rmd'}
```



# Model fitting

For all inversions, NH inversions and NAHR inversions, I have fitted an Ordinal Logistic Regression with `polr`, which returns a coefficient that represents the log(OddsRatio) and the corresponding Standard Error and t-value. For ease of interpretation, I have included in the table the Odds Ratio and the p-value corresponding to the t-value.

The summary tables also include the p-value for the Brant test, that checks whether the proportional odds assumption is true for this dataset (H0 = the proportional odds assumption holds). All models fulfill this assumption, although for the NAHR these p-values are to be taken with caution because some ChrRegion-invCategory combinations did not exist. 

The interpretation for the Odds Ratio is better understood with an example: given an independent variable x with an Odds Ratio of 1.25, for each increase in 1 measurement unit of variable x, a window is 1.25 times more likely to be in a higher inversion count category, given that the other variables remain constant. Since our variables are scaled, measurement units are standard deviations of each variable distribution. A significant p-value means that within the coefficient's confidence interval the values do not change sign, i.e. the variable is consistently increasing or decreasing the odds.

In the general model (Table 5) the variable Length is the only significant one, with an Odds Ratio = 2.197 and p-value = 0.002. When only NH inversions are taken into account  (Table 6), both the significance and the Odds Ratio increase: OR = 2.955 and p-value = 0.0002. On the other hand, when only NAHR inversions are considered (Table 7), the Length effect is no longer relevant and the only significant variable is Repeat number, with OR = 1.74 and p-value = 0.01. This confirms previous findings that at large window scales, the number of inversions depends on the generation mechanism, so the amount of repeats will be a key determinant of the amount of NAHR inversions in a region, while NH inversions will be generated randomly anywhere in the genome. It also evidences that when studying all inversions together, the patterns will be a mix of both inversion types, thus the importance of studying them separately given their very distinct behaviors. 

<!-- ## All inversions -->

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb.Scaled + allRepCounts.Scaled + ChrRegion + WAvgRate.perMb.Scaled"

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r , child = '03invAnalysisShort.Rmd'}
```



<!-- ## NH inversions -->

```{r, include=FALSE}
# Inversion type
yVar<-"NHCategory"

# Scaled/notScaled
xVars<- "Length.Mb.Scaled + allRepCounts.Scaled + ChrRegion + WAvgRate.perMb.Scaled"

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```




<!-- ## NAHR inversions -->

```{r, include=FALSE}
# Inversion type
yVar<-"NAHRCategory"

# Scaled/notScaled
xVars<- "Length.Mb.Scaled + allRepCounts.Scaled + ChrRegion + WAvgRate.perMb.Scaled"

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric", "Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

# Possible improvements / To do list

It would be nice to calculate power and goodness of fit for these models, especially the NAHR one, which is the one that needs more caution on interpretation. 