---
title: "Ordinal logistic model on large, classified windows data"
author: "Ruth Gómez Graciani"
output:
  pdf_document: default
header-includes:
 \usepackage{float}
 \floatplacement{figure}{H}
---

\tableofcontents

# Prepare the data

```{r child = '01PrepareTheData_CEUSpence.Rmd'}
```


# Descriptive statistics

```{r child = '02ExploreTheData_numCategories.Rmd'}
```


# Basic inversions model

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, telomeric, arm","Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

\newpage
# All dataset together without colors category

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,4), names=c(NA,"Low, Med-low, Med-high, High","Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

\newpage
# Centromeres vs telomeres models



## Centromeres & arms

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm","Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
load(file= "data/winRegions.RData")

winRegions<-winRegions[winRegions$Color != "telomeric",]
levels(winRegions$Color)<-c("arm", "centromeric", NA)
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

\newpage

## Arms

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,4), names=c(NA,"Low, Med-low, Med-high, High", "Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
load(file= "data/winRegions.RData")

winRegions<-winRegions[winRegions$Color == "arm",]
levels(winRegions$Color)<-c("arm", NA, NA)
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```

```{r child = '03invAnalysisShort.Rmd'}
```

\newpage
## Telomeres

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,4), names=c(NA,"Low, Med-low, Med-high, High", "Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
load(file= "data/winRegions.RData")

winRegions<-winRegions[winRegions$Color == "telomeric",]
levels(winRegions$Color)<-c(NA, NA, "telomeric")
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```

```{r child = '03invAnalysisShort.Rmd'}
```

# Without some values

## Without 3+
```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric","Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
load(file= "data/winRegions.RData")

winRegions<-winRegions[winRegions$invCategory != "3+",]
levels(winRegions$invCategory)<-c("0", "1", "2", NA)

save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

## Without 2

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric","Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
load(file= "data/winRegions.RData")

winRegions<-winRegions[winRegions$invCategory != "2",]
levels(winRegions$invCategory)<-c("0", "1", NA, "3+")

save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

## Without 1

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric","Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
load(file= "data/winRegions.RData")

winRegions<-winRegions[winRegions$invCategory != "1",]
levels(winRegions$invCategory)<-c("0", NA, "2", "3+")

save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

## Without 0

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, arm, telomeric","Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
load(file= "data/winRegions.RData")

winRegions<-winRegions[winRegions$invCategory != "0",]
levels(winRegions$invCategory)<-c(NA, "1", "2", "3+")

save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```


# Bingo! 1 categories are doing something...
```{r}
load(file= "data/winRegions.RData")


ggplot(winRegions)+geom_histogram(aes(x = WAvgRate.perMb))+
  facet_grid(invCategory ~ Color)

ggplot(winRegions, aes(x = "recRate", y = WAvgRate.perMb))+
      # Half violin
      ggdist::stat_halfeye(adjust = .5, width = .6, .width = 0, justification = -.2, point_colour = NA) +
      # Boxplot 
      geom_boxplot(width = .1, outlier.shape = NA) +
      # Points
      gghalves::geom_half_point_panel(side = "l", range_scale = .6,  alpha = .5, aes(color = invCategory))+
      # scale_color_manual(values = c(rep("#3c7ae7",11),rep("#89b23e",11) ))+
      # Adjust coordinates
      coord_flip()+facet_grid(Color ~ .)+
      # coord_flip( xlim = c(1.3, NA))+
      # Adjust labels
      theme(axis.title.y = element_blank(), legend.position = "top")+
      # Title
      ggtitle("group")


ggplot(winRegions, aes(x = "recRate", y = WAvgRate.perMb))+
      # Half violin
      ggdist::stat_halfeye(adjust = .5, width = .6, .width = 0, justification = -.2, point_colour = NA) +
      # Boxplot 
      geom_boxplot(width = .1, outlier.shape = NA) +
      # Points
      gghalves::geom_half_point_panel(side = "l", range_scale = .6,  alpha = .5, aes(color = Color))+
      # scale_color_manual(values = c(rep("#3c7ae7",11),rep("#89b23e",11) ))+
      # Adjust coordinates
      coord_flip()+facet_grid(invCategory ~ .)+
      # coord_flip( xlim = c(1.3, NA))+
      # Adjust labels
      theme(axis.title.y = element_blank(), legend.position = "top")+
      # Title
      ggtitle("group")
```

I will repeat the analysis without outliers to look for influential points.


## Without per-region (Color) outliers

```{r}
load(file= "data/winRegions.RData")


outvalue<-boxplot(winRegions$WAvgRate.perMb ~ winRegions$Color)$out

winRegions<-winRegions[!(winRegions$WAvgRate.perMb %in% outvalue),]
```


```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, telomeric, arm","Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

\newpage


## Without per-category (InvCategory) outliers

```{r}
load(file= "data/winRegions.RData")


outvalue<-boxplot(winRegions$WAvgRate.perMb ~ winRegions$invCategory)$out

winRegions<-winRegions[!(winRegions$WAvgRate.perMb %in% outvalue),]
```


```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, telomeric, arm","Low, Med-low, Med-high, High"), stringsAsFactors = F )

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

\newpage

## Without all outliers
```{r}
load(file= "data/winRegions.RData")


outvalue_A<-boxplot(winRegions$WAvgRate.perMb ~ winRegions$invCategory, horizontal = TRUE)$out
outvalue_B<-boxplot(winRegions$WAvgRate.perMb ~ winRegions$Color, horizontal = TRUE)$out

winRegions<-winRegions[!(winRegions$WAvgRate.perMb %in% outvalue_A),]
winRegions<-winRegions[!(winRegions$WAvgRate.perMb %in% outvalue_B),]

save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")

```

```{r child = '03invAnalysisShort.Rmd'}
```

## Manually remove group 1 outlier
```{r}
load(file= "data/winRegions.RData")


outvalue<-max(winRegions[winRegions$invCategory == "1", "WAvgRate.perMb"])

winRegions<-winRegions[!(winRegions$WAvgRate.perMb %in% outvalue),]
winRegions[winRegions$WAvgRate.perMb == outvalue,]
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")

```

```{r child = '03invAnalysisShort.Rmd'}
```

# Which is this point?

```{r}


load(file= "data/winRegions.RData")

# Setup
pacman::p_load(ggplot2)
name <-"CEUSpence_COzones_0.03_1000000"
divChroms <- paste0("data/divideChromosomes/",name,"/")

  # PATHS
  windowsFile <- paste0(divChroms,"/windows.txt")
  densityFile <- paste0(divChroms,"/densities.txt")
  extremesFile <- paste0(divChroms,"/extremes.txt")
  centroFile <- paste0(divChroms,"/workspace.txt")
  
  # FILES
  windows <- read.table(windowsFile, header = T, sep = "\t")
    windows$Color <- "a"
    windows[as.numeric(rownames(windows)) %% 2 == 1,"Color"]<-"b"
  windows$Chromosome <- factor(windows$Chromosome, levels =  paste(rep("chr", 23), as.character(c(c(1:22),"X")), sep = ""))
    
  density <- read.table(densityFile, header = T, sep = "\t")
  density$Chromosome <- factor(density$Chromosome, levels = paste(rep("chr", 23), as.character(c(c(1:22),"X")), sep = ""))
  
  extremes <- read.table(extremesFile, header = T, sep = "\t")
  extremes$Chromosome <- factor(extremes$Chromosome, levels = paste(rep("chr", 23), as.character(c(c(1:22),"X")), sep = ""))
  
  centromeres <- read.table(centroFile, header = T, sep = "\t")
  centromeres$Chromosome <- factor(centromeres$Chromosome, levels =  paste(rep("chr", 23), as.character(c(c(1:22),"X")), sep = ""))
  armLimits <-centromeres[grep("cen", centromeres$chromID,invert = T),]
  centromeres<-centromeres[grep("cen", centromeres$chromID),]
  starts <-  armLimits[grep("p", armLimits$chromID), c("Start", "Chromosome")]
  ends<- armLimits[grep("q", armLimits$chromID), c("End", "Chromosome")]
  
    d_rowlist<-c()
  c_rowlist<-c()
  for (c in unique(starts$Chromosome)) {
    cStart <-min(extremes[extremes$Chromosome ==c, "pos"])
    starts<- starts[ !(starts$Chromosome == c & starts$Start < cStart), ] # chromosome start will be first minimum
    windows[windows$Chromosome ==c & windows$Start < cStart,"Start" ]<-cStart #window start may be affected
    d_rowlist<-c(d_rowlist,rownames(density[density$Chromosome ==c & density$pos < cStart,])) #some densities will disappear
    c_rowlist<-c(c_rowlist,rownames(centromeres[centromeres$Chromosome ==c & centromeres$Start < cStart,])) # if starts with centromere will disappear


  }
  density<-density[!(rownames(density) %in% d_rowlist),]
  centromeres<-centromeres[!(rownames(centromeres) %in% c_rowlist),]
  
  # PLOT
  ggplot()+
        geom_rect(data= centromeres,  aes(xmin = Start, xmax = End,  ymin = 0, ymax = Inf), fill = "blue4", alpha = 0.5)+
        geom_vline(data=starts, aes(xintercept = Start), color = "orange")+geom_vline(data=ends, aes(xintercept = End), color = "orange")+
        geom_rect(data=windows, aes(xmin = Start, xmax = End, fill = Color, ymin = 0, ymax = Inf), alpha = 0.3)+
        geom_line(data=density,aes(x = pos, y = val))+
        geom_point(data=extremes, aes(x = pos, y = val, color = Type))+
        facet_wrap("Chromosome", scales  = "free", ncol=4)+
        scale_fill_manual(values=c("#737373", "#e1e5eb"), guide="none")+
        scale_color_manual(values = c("#bd2b43", "#2b63bd"), guide = "none")+
        scale_y_continuous(labels=function(x)x*1000000)+
         theme(axis.ticks.x = element_blank(),axis.text.x = element_blank(),axis.text.y = element_text(size = 5))+
        ylab("Rate (cM/Mb)")+xlab("Position")+ggtitle(name) 


  ggplot()+
        # geom_rect(data= centromeres,  aes(xmin = Start, xmax = End,  ymin = 0, ymax = Inf), fill = "blue4", alpha = 0.5)+
        geom_vline(data=starts, aes(xintercept = Start), color = "orange")+geom_vline(data=ends, aes(xintercept = End), color = "orange")+
        geom_rect(data=winRegions, aes(xmin = Start, xmax = End, fill = Color, ymin = 0, ymax = WAvgRate.perMb), alpha = 0.3)+
        geom_line(data=density,aes(x = pos, y = val * 1000000 ))+
        geom_text(data = winRegions, aes(x = (Start+End)/2 , y = WAvgRate.perMb, label = invCategory))+
        geom_point(data=extremes, aes(x = pos, y =val * 1000000 , color = Type))+
        facet_grid("Chromosome", scales = "free_x", space = "free_x")+
        # scale_fill_manual(values=c("#737373", "#e1e5eb"), guide="none")+
        scale_color_manual(values = c("#bd2b43", "#2b63bd"), guide = "none")+
        # scale_y_continuous(labels=function(x)x*1000000)+
         theme(axis.ticks.x = element_blank(),axis.text.x = element_blank(),axis.text.y = element_text(size = 5), legend.position = "top")+
        ylab("Rate (cM/Mb)")+xlab("Position")+ggtitle(name) 
  
```

# Wihtout some chromosomes


```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, telomeric, arm","Low, Med-low, Med-high, High"), stringsAsFactors = F )

load(file= "data/winRegions.RData")

winRegions<-winRegions[winRegions$Chromosome %in% c("chr6", "chr7", "chr8", "chr9", "chr10"),]
# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

Tested unsuccessfully: 

- no Chr22
- no Chr21
- no Chr19
- chroms 1,2,3,4,5
- chroms 6,7,8,9,10

So it seems something about the data...

# Multiply the numbers

## cM/Mb *10
```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, telomeric, arm","Low, Med-low, Med-high, High"), stringsAsFactors = F )

load(file= "data/winRegions.RData")

winRegions$WAvgRate.perMb<-winRegions$WAvgRate.perMb*10
# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

## cM/Mb *100

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, telomeric, arm","Low, Med-low, Med-high, High"), stringsAsFactors = F )

load(file= "data/winRegions.RData")

winRegions$WAvgRate.perMb<-winRegions$WAvgRate.perMb*100
# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```
## cM/Mb *1000


```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, telomeric, arm","Low, Med-low, Med-high, High"), stringsAsFactors = F )

load(file= "data/winRegions.RData")

winRegions$WAvgRate.perMb<-winRegions$WAvgRate.perMb*1000
# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

## cM/Mb *10000
```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, telomeric, arm","Low, Med-low, Med-high, High"), stringsAsFactors = F )

load(file= "data/winRegions.RData")

winRegions$WAvgRate.perMb<-winRegions$WAvgRate.perMb*10000
# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

## cM/Mb x 10000 & Length x 10000

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, telomeric, arm","Low, Med-low, Med-high, High"), stringsAsFactors = F )

load(file= "data/winRegions.RData")

winRegions$WAvgRate.perMb<-winRegions$WAvgRate.perMb*10000
winRegions$Length.Mb<-winRegions$Length.Mb/10000
# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}
```

## Only centimorgans

```{r, include=FALSE}
# Inversion type
yVar<-"invCategory"

# Scaled/notScaled
xVars<- "Length.Mb + allRepCounts + Color + WAvgRate.perMb" 

# Formula
myFormula<-paste0(yVar,"~", xVars)
xVarsVector<-strsplit(xVars, " [+] ")[[1]]

# Samples for prediction
samples<-data.frame(vars=xVarsVector, sam=c(50,4,0,4), names=c(NA,"Low, Med-low, Med-high, High", "centromeric, telomeric, arm","Low, Med-low, Med-high, High"), stringsAsFactors = F )

load(file= "data/winRegions.RData")

cor(winRegions$WAvgRate.perMb, winRegions$Length.Mb)
plot(scale(winRegions$WAvgRate.perMb), scale(winRegions$Length.Mb))

# Pass to file
save(yVar,xVars,myFormula,xVarsVector,samples , winRegions, file= "data/analysisParams.RData")
```


```{r child = '03invAnalysisShort.Rmd'}



```


