<!-- # --- -->
<!-- # title: "Ordinal logistic model on large windows data" - chapter 1 -->
<!-- # author: "Ruth Gómez Graciani" -->
<!-- # output: -->
<!-- #   pdf_document: default -->
<!-- # header-includes: -->
<!-- #  \usepackage{float} -->
<!-- #  \floatplacement{figure}{H} -->
<!-- # --- -->

```{r, include=FALSE}
if (!require("pacman")) install.packages("pacman")
# set conditions for code output
knitr::opts_chunk$set(message=FALSE, warning=FALSE, error=FALSE,fig.pos = "!H")
showcode=FALSE #for include or echo - echo=showcode // include=showcode
runbash=FALSE # for eval=runbash
runR=TRUE #for eval=runR
```


# Prepare the data

First, we obtain the density distribution, and local minima and maxima for the recombination map. 

<!-- Bash run to make standard windows -->

```{bash, include=showcode, eval=runbash}
ENVPY="/home/rgomez/anaconda3/bin/python"
cd /home/rgomez/Documents/PHD_Repos/20211117_PopRecombination/

OUTDIR="report/20220519_tutorialLogisticModel/LargeClassifiedWindowsSpence/data"

$ENVPY code/python/divideChromosomes.py \
  -m data/SpenceSong_hg19_recMaps_processed/CEU_recombination_map_hg19_allChr.bed  \
  -b data/cytoBand.txt -f 5 -s 0 -r 1000000 -o $OUTDIR -d CEUSpence


cd report/20220519_tutorialLogisticModel/LargeClassifiedWindowsSpence/
```

<!-- Windows visualization -->

```{r Base-workspace, eval=runR,echo=showcode, fig.cap="Crossover zones; centromeres in blue, workspace limits in orange." ,  fig.height=6}
# Setup
pacman::p_load(ggplot2)
name <-"CEUSpence_COzones_0.0_800000"
divChroms <- paste0("data/divideChromosomes/",name,"/")

  # PATHS
  windowsFile <- paste0(divChroms,"/windows.txt")
  densityFile <- paste0(divChroms,"/densities.txt")
  extremesFile <- paste0(divChroms,"/extremes.txt")
  centroFile <- paste0(divChroms,"/workspace.txt")
  
  # FILES
  windows <- read.table(windowsFile, header = T, sep = "\t")
    windows$Color <- "a"
    windows[as.numeric(rownames(windows)) %% 2 == 1,"Color"]<-"b"
  windows$Chromosome <- factor(windows$Chromosome, levels =  paste(rep("chr", 23), as.character(c(c(1:22),"X")), sep = ""))
    
  density <- read.table(densityFile, header = T, sep = "\t")
  density$Chromosome <- factor(density$Chromosome, levels = paste(rep("chr", 23), as.character(c(c(1:22),"X")), sep = ""))
  
  extremes <- read.table(extremesFile, header = T, sep = "\t")
  extremes$Chromosome <- factor(extremes$Chromosome, levels = paste(rep("chr", 23), as.character(c(c(1:22),"X")), sep = ""))
  
  centromeres <- read.table(centroFile, header = T, sep = "\t")
  centromeres$Chromosome <- factor(centromeres$Chromosome, levels =  paste(rep("chr", 23), as.character(c(c(1:22),"X")), sep = ""))
  armLimits <-centromeres[grep("cen", centromeres$chromID,invert = T),]
  centromeres<-centromeres[grep("cen", centromeres$chromID),]
  starts <-  armLimits[grep("p", armLimits$chromID), c("Start", "Chromosome")]
  ends<- armLimits[grep("q", armLimits$chromID), c("End", "Chromosome")]
  
  # Remove chr p arm data when necessary
  #### I will manyally remove the extremes before chr21 centromere
  out<- rownames(extremes[(extremes$Chromosome == "chr21" & extremes$pos < centromeres[centromeres$Chromosome == "chr21", "Start"]), ])
  extremes<-extremes[-as.numeric(out),]
  
  d_rowlist<-c()
  c_rowlist<-c()
  for (c in unique(starts$Chromosome)) {
    cStart <-min(extremes[extremes$Chromosome ==c, "pos"])
    starts<- starts[ !(starts$Chromosome == c & starts$Start < cStart), ] # chromosome start will be first minimum
    windows[windows$Chromosome ==c & windows$Start < cStart,"Start" ]<-cStart #window start may be affected
    d_rowlist<-c(d_rowlist,rownames(density[density$Chromosome ==c & density$pos < cStart,])) #some densities will disappear
    c_rowlist<-c(c_rowlist,rownames(centromeres[centromeres$Chromosome ==c & centromeres$Start < cStart,])) # if starts with centromere will disappear


  }
  density<-density[!(rownames(density) %in% d_rowlist),]
  centromeres<-centromeres[!(rownames(centromeres) %in% c_rowlist),]

  
  # PLOT
  (p<-ggplot()+
        geom_rect(data= centromeres,  aes(xmin = Start, xmax = End,  ymin = 0, ymax = Inf), fill = "blue4", alpha = 0.5)+
        geom_vline(data=starts, aes(xintercept = Start), color = "orange")+geom_vline(data=ends, aes(xintercept = End), color = "orange")+
        geom_rect(data=windows, aes(xmin = Start, xmax = End, fill = Color, ymin = 0, ymax = Inf), alpha = 0.3)+
        geom_line(data=density,aes(x = pos, y = val))+
        geom_point(data=extremes, aes(x = pos, y = val, color = Type))+
        facet_wrap("Chromosome", scales  = "free", ncol=4)+
        scale_fill_manual(values=c("#737373", "#e1e5eb"), guide="none")+
        scale_color_manual(values = c("#bd2b43", "#2b63bd"), guide = "none")+
        scale_y_continuous(labels=function(x)x*1000000)+
         theme(axis.ticks.x = element_blank(),axis.text.x = element_blank(),axis.text.y = element_text(size = 5))+
        ylab("Rate (cM/Mb)")+xlab("Position")+ggtitle(name) 
  )

  
  # invs<-read.table("data/invs.bed")
  # colnames(invs)<-c("Chromosome", "Start", "End", "Inv")
  # 
 # p+geom_point(data=invs, aes(x=(End+Start)/2, y = 0.004/1000000))
```

\newpage

Next, we define telomeric regions as the space between the chromosome start to the next local minimum, or between the chromosome end to the previous local minimum. We also define the limits of the centromere by calculating the midpoint between the flanking maxima and the local minimum near the centromere (or the centromere itself if there is no local minimum). Actual centromeres will be excluded from centromeric regions to avoid biases, specially in the Spence data. These categories will be represented as the "Color" variable in the statistical analysis. 

```{r Modified-workspace, eval=runR, echo=showcode,  fig.cap="Color-coded windows for telomeric, centromeric and arm categories.",  fig.height=6 }
windows <- data.frame()

for (chrom in unique(extremes$Chromosome)) {
  
  # Sort extrema
  extremes_c<- extremes[which(extremes$Chromosome == chrom) ,]
  extremes_c<-extremes_c[order(extremes_c$pos),]
  
  # Chromosome extremes
  teloExtremes<-extremes_c$pos[c(1,nrow(extremes_c))]
  
  # Telomeres threshold option A (midpoint between Max and Min)
  teloMidsA<-c((extremes_c$pos[2]+extremes_c$pos[3])/2, (extremes_c$pos[nrow(extremes_c)-1]+extremes_c$pos[nrow(extremes_c)-2])/2)
  telomeric<-data.frame(pos=c(teloExtremes, teloMidsA), Chromosome = chrom)
  telomeric<-telomeric[order(telomeric$pos),]
  
  # Centromeres
  centromeric <- centromeres[centromeres$Chromosome == chrom,]
  if (nrow(centromeric) != 0){
    leftMax<-max(extremes_c[which(extremes_c$pos <= centromeric$Start & extremes_c$Type == "Maxima"), "pos"])
    rightMax<-min(extremes_c[which(extremes_c$pos >= centromeric$End & extremes_c$Type == "Maxima"), "pos"])
    
    centromeric$Start<- (centromeric$Start + leftMax)/2
    centromeric$End<-(centromeric$End + rightMax)/2
  }
  
  # Overlaps
  if (nrow(centromeric) != 0){
    #Average centromeres
    teloend<- telomeric$pos[2]
    telomeric$pos[2]<-ifelse(teloend>=centromeric$Start,  centromeric$Start , teloend)
    telostart<-telomeric$pos[3]
    telomeric$pos[3]<-ifelse(telostart<=centromeric$End, centromeric$End, telostart )
  }

  #Positions
  positions<-unique(c(telomeric$pos, centromeric$Start, centromeric$End))
  positions<-positions[order(positions)]
  
  chromwins<-data.frame(Start = positions[1:length(positions)-1], End = positions[2:length(positions)], Chromosome = chrom)
  chromwins$Color<- "arm"
  if (nrow(centromeric) != 0){
    chromwins[chromwins$Start %in% telomeric[c(1,3), "pos"],"Color"]<-"telomeric"
    chromwins[chromwins$Start == centromeric$Start,"Color"]<-"centromeric"
    
    # Divide centromeric in two
    cenBis<-chromwins[chromwins$Color == "centromeric",]

    chromwins[chromwins$Color == "centromeric","End"] <- centromeres[centromeres$Chromosome == chrom,"Start"]
    cenBis$Start <- centromeres[centromeres$Chromosome == chrom,"End"]

    chromwins<-rbind(chromwins, cenBis)

  }else{
    chromwins[chromwins$Start %in% telomeric[c(3), "pos"],"Color"]<-"telomeric"
  }
  

  windows <- rbind(windows,chromwins)
}


write.table(windows, "data/windows.txt", quote = F, row.names = F, col.names = T, sep = "\t")



  ggplot()+
        # geom_rect(data= centromeres,  aes(xmin = Start, xmax = End,  ymin = 0, ymax = Inf), fill = "blue4", alpha = 0.5)+
        geom_vline(data=starts, aes(xintercept = Start), color = "orange")+geom_vline(data=ends, aes(xintercept = End), color = "orange")+
        geom_rect(data=windows, aes(xmin = Start, xmax = End, fill = Color, ymin = 0, ymax = Inf), alpha = 0.3)+
        geom_line(data=density,aes(x = pos, y = val))+
        geom_point(data=extremes, aes(x = pos, y = val, color = Type))+
        facet_wrap("Chromosome", scales = "free", ncol = 4)+
        # scale_fill_manual(values=c("#737373", "#e1e5eb"), guide="none")+
        scale_color_manual(values = c("#bd2b43", "#2b63bd"), guide = "none")+
        scale_y_continuous(labels=function(x)x*1000000)+
         theme(axis.ticks.x = element_blank(),axis.text.x = element_blank(),axis.text.y = element_text(size = 5), legend.position = "top")+
        ylab("Rate (cM/Mb)")+xlab("Position")+ggtitle(name) 
  

```

<!-- Fill window information -->

```{bash, include=showcode, eval=runbash}
ENVPY="/home/rgomez/anaconda3/bin/python"
NAME="LargeClassifiedWindowsSpence"
cd "../../../"
$ENVPY code/python/fillWindowMeasurements.py -m data/SpenceSong_hg19_recMaps_processed/CEU_recombination_map_hg19_allChr.bed -w report/20220519_tutorialLogisticModel/${NAME}/data/windows.txt  -o report/20220519_tutorialLogisticModel/${NAME}/data/ -d "chromRegions"
cd report/20220519_tutorialLogisticModel/${NAME}/
```
